% ==============================================================================
% CEC 320 FINAL EXAM CHEATSHEET --- PART 1 PREAMBLE
% Auto-generated header. Concatenate with partials in part1/ then _footer.tex.
% Built by build.sh.
% ==============================================================================
\documentclass[8pt,landscape]{extarticle}
\usepackage[letterpaper,landscape,margin=0.18in,top=0.18in,bottom=0.18in]{geometry}
\usepackage{multicol}
\usepackage{amsmath,amssymb}
\usepackage{array}
\usepackage{booktabs}
\usepackage{xcolor}
\usepackage{listings}
\usepackage{enumitem}
\usepackage{titlesec}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{microtype}

\AtBeginDocument{\small}

% ---------- Spacing / layout ----------
\setlength{\columnsep}{0.3em}
\setlength{\columnseprule}{0.15pt}
\setlength{\parindent}{0pt}
\setlength{\parskip}{1pt}
\setlist{nosep,leftmargin=0.9em,itemsep=0pt,topsep=0.5pt,parsep=0pt,partopsep=0pt}
\setlength{\tabcolsep}{2pt}
\renewcommand{\arraystretch}{1.0}
\pagestyle{empty}
\raggedcolumns

% ---------- Section styling (tight) ----------
\titleformat{\section}{\normalfont\bfseries\footnotesize\color{blue!60!black}}
  {\thesection.}{0.25em}{}
\titlespacing*{\section}{0pt}{3pt}{0.5pt}
\titleformat{\subsection}{\normalfont\bfseries\scriptsize}{}{0em}{}
\titlespacing*{\subsection}{0pt}{2pt}{0.2pt}

% ---------- ARM listings ----------
\lstdefinelanguage{arm}{
  morekeywords={ldr,ldrb,ldrh,ldrsb,ldrsh,ldrd,str,strb,strh,strd,mov,movw,movt,%
    movs,add,adds,adc,sub,subs,sbc,cmp,cmn,tst,teq,bl,bx,b,beq,bne,bgt,bge,blt,%
    ble,bhi,bhs,blo,bls,bcs,bcc,push,pop,cbz,cbnz,it,itt,ite,itte,itet,itee,%
    ittt,ittee,iteee,ittte,itttt,tbb,tbh,adr,lsl,lsls,lsr,lsrs,asr,asrs,ror,%
    rors,rrx,rrxs,and,ands,orr,orrs,eor,eors,bic,bics,orn,mvn,mvns,bfc,bfi,%
    mul,muls,mla,mls,umull,umlal,smull,smlal,udiv,sdiv,nop,wfi,wfe,svc,rsb,%
    sxtb,sxth,uxtb,uxth,clz,rbit,rev,%
    lslge,lslgt,lsllt,lslle,asrge,asrgt,asrlt,asrle,ldrle,ldrgt,ldrlt,ldrge,%
    ldrshgt,ldrshlt,ldrhge,ldrhlt,lsrlt,lsrge},
  morecomment=[l]{@},
  sensitive=false
}
\lstset{
  language=arm,
  basicstyle=\ttfamily\scriptsize,
  keywordstyle=\color{blue!70!black}\bfseries,
  commentstyle=\color{green!40!black}\itshape,
  breaklines=false,
  frame=none,
  aboveskip=0.5pt,
  belowskip=0.5pt,
  columns=fullflexible,
  keepspaces=true,
  xleftmargin=0.25em,
  showstringspaces=false
}
\newcommand{\asm}[1]{\texttt{\scriptsize #1}}
\newcommand{\cee}[1]{\texttt{\scriptsize #1}}

% Tag environments
\newenvironment{ex}
  {\par\smallskip\noindent{\scriptsize\color{red!55!black}\textbf{Ex.}}\,\scriptsize\ignorespaces}
  {\par}
\newenvironment{qp}
  {\par\smallskip\noindent{\scriptsize\color{purple!70!black}\textbf{Qz.}}\,\scriptsize\ignorespaces}
  {\par}
\newenvironment{hwp}
  {\par\smallskip\noindent{\scriptsize\color{orange!70!black}\textbf{HW}}\,\scriptsize\ignorespaces}
  {\par}
\newenvironment{reg}
  {\par\smallskip\noindent{\scriptsize\color{teal!75!black}\textbf{Reg}}\,\scriptsize\ignorespaces}
  {\par}

\begin{document}
\begin{center}
  \textbf{\footnotesize CEC 320 --- FINAL EXAM Cheatsheet (Part 1)} \;
  \scriptsize HW 5--12 \textbullet{} Lctr 14--27 \textbullet{} Quiz 3, 4, 6a
\end{center}
\vspace{-8pt}

\begin{multicols*}{4}
\scriptsize

% ----- 01_lectures_14_27.tex -----
\section{Lctr 14--27 Recall Bullets}

\subsection{L14 --- Mixed C/ASM (EABI)}
\begin{itemize}
\item EABI: \asm{R0--R3} pass args + return; caller-saved (NOT preserved). \asm{R4--R11} callee-saved $\Rightarrow$ push/pop if used.
\item 64-bit arg/return: low word in \asm{R0}, high word in \asm{R1} (little-endian).
\item Stack = Full Descending: SP points at last pushed item; \asm{push} decrements SP by 4/word.
\item Multi-reg push: low-numbered regs end up at LOW addresses regardless of brace order in \asm{\{...\}}.
\item Return: \asm{bx lr} if LR not stacked; \asm{pop \{pc\}} if LR was pushed ($\equiv$\asm{pop \{lr\}}+\asm{bx lr}).
\item Leaf fn $\Rightarrow$ no LR push needed; stem fn (calls another) MUST push LR before BL.
\end{itemize}

\subsection{L15 --- Arithmetic}
\begin{itemize}
\item \asm{ADD/SUB/RSB} (no S) leave flags alone; \asm{ADDS/SUBS/RSBS} update NZCV. \asm{RSB} exists because Rn must be a reg.
\item \asm{ADC Rn,Op2}$=$Rn$+$Op2$+$C; \asm{SBC Rn,Op2}$=$Rn$-$Op2$+$C$-$1 (C$=$1 means no-borrow).
\item 64-bit add idiom: \asm{adds r0,r2} / \asm{adc r1,r3} (low sets C, high consumes it).
\item \asm{MUL/MLA/MLS} 32-bit signed$\equiv$unsigned in low 32; \asm{UMULL/SMULL/UMLAL/SMLAL} for 64-bit.
\item \asm{UDIV} vs \asm{SDIV} differ; no flag-setting form, non-constant cycles.
\item Modulo idiom: \asm{udiv r2,r0,r1} / \asm{mls r0,r1,r2,r0} ($r0=r0-r1\cdot r2$).
\end{itemize}

\subsection{L16 --- Shifts \& Op2}
\begin{itemize}
\item \asm{LSL} both signed/unsigned; \asm{LSR} unsigned-only (0-fill); \asm{ASR} signed-only (sign-bit replicate). Last bit out $\to$ C.
\item \asm{RRX} rotates right by 1 through C: C $\to$ MSB, LSB $\to$ C.
\item Op2 forms: \asm{\#imm}, \asm{Rm}, or \asm{Rm,LSL/LSR/ASR \#n} (barrel shifter, free combined w/ arithmetic).
\item Shifted-reg Op2: destination canNOT be omitted even if same as Rn.
\item Q31 accurate mul (3 instr): \asm{smull r1,r0,r0,r1} / \asm{lsl r0,\#1} / \asm{add r0,r0,r1,lsr \#31}.
\item Fast scaling by $(2^n\pm1)/2^n$: \asm{add r0,r0,r0,asr \#2}$=5/4$; \asm{sub r0,r0,r0,asr \#3}$=7/8$.
\end{itemize}

\subsection{L17 --- Bitwise Logic}
\begin{itemize}
\item \asm{MVN}$=\sim$, \asm{AND}$=\&$, \asm{ORR}$=|$, \asm{EOR}$=\hat{}\,$. No-C-equivalents: \asm{BIC Rn,Op2}$=$Rn\,\&\,$\sim$Op2; \asm{ORN}$=$Rn\,$|$\,$\sim$Op2.
\item \asm{BFC Rd,\#s,\#w} clears $w$ bits at pos $s$; \asm{BFI Rd,Rn,\#s,\#w} inserts low-$w$ bits of Rn at pos $s$.
\item Clear-bit-N: \asm{mov mask,\#1}/\asm{lsl mask,\#N}/\asm{bic r0,mask} (BIC$=$AND-NOT).
\item Read APSR: \asm{MRS r0,apsr}; write: \asm{MSR apsr\_nzcvq,Rn}.
\item Bit-field assign $v$: build mask $((1\!\ll\!w)-1)\!\ll\!s$, \asm{bic} target, shift $v$ to pos, \asm{and} mask, \asm{orr} in.
\end{itemize}

\subsection{L18 --- NZCV Flags}
\begin{itemize}
\item \textbf{N}$=$MSB(result); \textbf{Z}$=$(result$=$0); \textbf{C}$=$unsigned carry/no-borrow; \textbf{V}$=$signed overflow.
\item HW always updates BOTH C and V; programmer chooses interpretation.
\item Unsigned uses \{C,N,Z\}; signed uses \{V,N,Z\}.
\item C on SUB: C$=$1 if no borrow (Rn$\ge$Op2), C$=$0 if borrow (counterintuitive naming).
\item \asm{CMP}$=$SUB w/o writeback (sets flags). \asm{CMN}$=$add-form. \asm{TST}$=$AND-only-flags. \asm{TEQ}$=$EOR-only-flags.
\item N-bit verify trick: \asm{LSL \#(32-N)} operands, \asm{ADDS}, \asm{ASR \#(32-N)} back, read APSR via \asm{MRS}.
\end{itemize}

\subsection{L19 --- CCS \& Conditional Branch}
\begin{itemize}
\item Signed CCS: \asm{EQ/NE/GT/GE/LT/LE}. Unsigned: \asm{EQ/NE/HI/HS(CS)/LO(CC)/LS}.
\item \asm{HS}$=$\asm{CS} (C set $\Rightarrow$ unsigned $\ge$); \asm{LO}$=$\asm{CC} (C clear $\Rightarrow$ unsigned $<$).
\item Mnemonic ``HiLow unSun'': \textbf{Hi/Low} family $=$ unSigned; \textbf{G/L} ($>$/$<$) $=$ Signed.
\item Consistency rule: signed operands $\Rightarrow$ signed CCS + \asm{ASR} for divide; unsigned $\Rightarrow$ unsigned CCS + \asm{LSR}.
\item \asm{cmp r0,\#-5} assembled as \asm{CMN r0,\#5}; signed CCSs still behave correctly.
\end{itemize}

\subsection{L20 --- LDR/STR Immediate}
\begin{itemize}
\item Three forms: \asm{[Ra,\#os]} basic (no update); \asm{[Ra,\#os]!} pre (update BEFORE, ``urgent bang''); \asm{[Ra],\#os} post (update AFTER).
\item LOAD size: none(word), \asm{B}(u8), \asm{SB}(i8), \asm{H}(u16), \asm{SH}(i16). STORE: NO sign variant (truncate only).
\item C$\to$ASM: \asm{*p++}$=$post-index; \asm{*++p}$=$pre-index; \asm{*(p+k)}$=$basic offset, $k\cdot$sizeof(elem).
\item \asm{STRB} writes only low byte of Rv; upper bits ignored (no signed variant needed).
\item Valid offset range: word LDR/STR typically $\pm$4095; byte/halfword $\pm$255 on certain forms.
\end{itemize}

\subsection{L21 --- LDR/STR Register + MOV}
\begin{itemize}
\item Reg-offset: \asm{LDR Rv,[Ra,Ri,LSL \#n]}, $n=\log_2$(elem-size): word$=$2, half$=$1, byte$=$0, dword$=$3.
\item NO writeback form for reg-offset (only immediate-offset has \asm{!} or trailing \asm{,\#os}).
\item Iter idioms: post-index advances Ra; reg-offset keeps Ra constant, increments Ri.
\item \asm{MOV} $=$ small imm only; \asm{MOVW} loads 16b low; \asm{MOVT} loads 16b high (upper halfword).
\item Pseudo \asm{LDR Rd,=expr} $\to$ assembler picks \asm{MOV}/\asm{MOVW}+\asm{MOVT}/PC-relative literal-pool load.
\end{itemize}

\subsection{L22 --- Increment Ops (vars + ptrs)}
\begin{itemize}
\item C precedence: postfix \asm{++} level 1 (LTR, highest); prefix \asm{++},\asm{*},\asm{\&}, cast all level 2 (RTL).
\item \asm{*p++} $=$ deref OLD p, then p advances. \asm{*++p} $=$ p advances, then deref NEW.
\item \asm{++*p} $=$ load val, +1, store back (modifies \emph{value}, not pointer).
\item \asm{++*p++}: post-inc p (so p now points past), but modify mem at OLD addr via \asm{[r0,\#-1]}.
\item Ptr arith scales by sizeof(elem): \asm{p8++}$+1$, \asm{p16++}$+2$, \asm{p32++}$+4$. Diff of typed ptrs $\to$ elements.
\item \asm{B[i]=A[i++]++;} is UB (unsequenced i mods); split or use comma operator.
\end{itemize}

\subsection{L23 --- ASM Calls ASM (LDRD/STRD)}
\begin{itemize}
\item \textbf{Leaf} fn (never BLs): \asm{bx lr} suffices, no LR push.
\item \textbf{Stem} fn (calls another): MUST \asm{push \{lr\}} (BL clobbers LR), \asm{pop \{pc\}} to return.
\item \asm{LDRD Rv\_L,Rv\_H,[Ra,\#os]}: 8 bytes, low word at low addr, high word at high addr (little-endian).
\item Rt2$=$Rt$+1$ (consecutive regs). NO register-offset form for \asm{LDRD/STRD} (basic/pre/post only).
\item Arg-prep by C type: int32$\to$\asm{LDR}; int16$\to$\asm{LDRSH}; uint16$\to$\asm{LDRH}; int8$\to$\asm{LDRSB}; uint8$\to$\asm{LDRB}; int64$\to$\asm{LDRD}.
\end{itemize}

\subsection{L24 --- Combo Branch / CEX (CBZ/IT)}
\begin{itemize}
\item \asm{CBZ Rn,label} / \asm{CBNZ Rn,label}: FORWARD only ($\sim$126 byte range), Rn $\in$ R0--R7, does NOT read/set flags.
\item Perfect after \asm{ldrb} for null-terminator test (loaded byte is flag-neutral).
\item IT block $=$ up to 4 conditional instr sharing one CCS. Forms: \asm{IT, ITT, ITE, ITTT, ITTE, ITEE, ITTEE, ITTTT}.
\item First IT slot is ALWAYS T (matches CCS); later slots T (same) or E (inverse). Each instr inside still needs explicit CCS suffix.
\item HARD bans: \asm{CBZ}/\asm{CBNZ} forbidden inside IT; conditional branch only allowed in LAST IT slot.
\end{itemize}

\subsection{L25 --- If / If-Else (PL/NL/CEX)}
\begin{itemize}
\item \textbf{PL} (positive logic): same-CCS branch to \asm{\_then}; else-block laid FIRST. Reads ``backward'' from C.
\item \textbf{NL} (negative logic): inverse-CCS branch to \asm{\_else}; then-block FIRST. Closer to C source order.
\item if-else-if: chain \asm{cmp}+NL-branch; \asm{b end} at each arm tail to skip the rest.
\item Compound \textbf{OR} (short-circuit): two CMPs each branch to shared \asm{\_then}; fall-through $=$ else.
\item Compound \textbf{AND} (De Morgan): both CMPs branch to \asm{\_else} on FAILURE; fall-through $=$ then.
\item CEX style: small bodies inside IT block instead of branches (saves a label).
\end{itemize}

\subsection{L26 --- Loops}
\begin{itemize}
\item Labels: \asm{<fn>\_lp} (top), \asm{<fn>\_end} (exit).
\item Top-tested: \asm{for}, \asm{while} (0+ iters). Bottom-tested: \asm{do-while} ($\ge$1 iter).
\item \asm{while} top-test form: NL-branch to \asm{\_end} at top, body, update, \asm{b lp}.
\item \asm{while} late-check: jump-over-body on entry, test at bottom $\Rightarrow$ saves one branch per iteration.
\item \asm{while} CEX form: \asm{cmp}+\asm{it cond}+conditional B as last slot in IT (only legal cond-branch-in-IT).
\item Priming-read pattern: \asm{ldrb} before loop + \asm{ldrb} inside loop (eliminated by \asm{while(1)+break}).
\end{itemize}

\subsection{L27 --- break / continue / switch}
\begin{itemize}
\item \asm{break} $=$ forward \asm{b <loop\_end>} (or \asm{CBZ}/\asm{CBNZ} to end on a value test).
\item \asm{continue} in while/do-while $\to$ branch to CONDITION test; in \asm{for} $\to$ branch to UPDATE expr (so \asm{i++} still runs!).
\item \asm{while(1) \{ ...; if(c) break; \}} removes priming-read duplication $\Rightarrow$ single \asm{ldrb} inside loop.
\item Switch naive: \asm{CMP}/\asm{BEQ} ladder; stacked case labels w/ no code between $\Rightarrow$ fall-through.
\item \asm{ADR Rd,label} $=$ PC-relative addr load (position-independent).
\item \asm{TBB [Rn,Rm]}: \asm{PC = PC+4+2*table[Rm]}; table bytes $=$ \asm{(case\_X-first\_case)/2} (Thumb-2 halfword aligned).
\end{itemize}

% ----- 02_hw05_qformat.tex -----
\section{HW5: Q-Format Fixed-Point Real Numbers}

\subsection{Format anatomy}
\textbf{Qm.n} (signed): 1 sign + $m$ int + $n$ frac $=$ $N=1+m+n$ bits, two's-comp.
\textbf{UQm.n} (unsigned): $m$ int + $n$ frac $=$ $N=m+n$ bits.
Scale factor $=2^{-n}$. LSB resolution $=2^{-n}$.

\begin{tabular}{@{}lllll@{}}
\toprule
Fmt & $N$ & $n$ & range (signed) & res \\
\midrule
Q3.4   & 8  & 4  & $-8.0\ldots+7.9375$ & $1/16$ \\
UQ3.5  & 8  & 5  & $0.0\ldots 7.96875$ & $1/32$ \\
Q2.5   & 8  & 5  & $-4.0\ldots+3.96875$ & $1/32$ \\
Q15 ($=$Q0.15) & 16 & 15 & $-1.0\ldots+0.99997$ & $2^{-15}$ \\
Q15.16 & 32 & 16 & $\pm 32{\rm k}$ & $2^{-16}$ \\
\bottomrule
\end{tabular}

\subsection{Encode rule (real $\to$ hex code)}
\begin{enumerate}
\item Multiply: $I = f\cdot 2^{n}$
\item Round to nearest integer.
\item If $I<0$ (signed Qm.n): $I \leftarrow I + 2^{N}$ (two's-comp wrap).
\item Express $I$ in hex (width $=\lceil N/4\rceil$ digits).
\end{enumerate}
Out-of-range $\Rightarrow$ saturate or flag overflow.

\subsection{Decode rule (hex code $\to$ real)}
\begin{enumerate}
\item Read $U =$ unsigned int value of code.
\item If signed Qm.n and MSB$=$1: $I = U - 2^{N}$; else $I=U$.
\item $f = I \cdot 2^{-n} = I/2^{n}$.
\end{enumerate}

\subsection{Multiplication / MAC rule}
$\text{Q}m.n \times \text{Q}p.q = \text{Q}(m{+}p{+}1).(n{+}q)$ in $2N$-bit product.
Renormalize back to Q$m.n$ by arithmetic shift right $n$ (\asm{ASR \#n}).
\textbf{Q15 MAC:} \asm{smull} gives Q30 in 64b across (HW:LW); shift the
\emph{low word} right by 15 to recover Q15 product. Sum of Q15 products is
plain integer add (no scaling).

\subsection{Q15.16 result extraction (32-bit Q15.16 from \texttt{smull})}
\asm{smull lo,hi,a,b} gives $a\cdot b$ as Q30.32 in (\asm{hi}:\asm{lo}).
To get Q15.16 (32b): \asm{lsl hi,\#16} (keep low 16 bits of hi as int part),
\asm{lsr lo,\#16} (keep high 16 bits of lo as frac part), then
\asm{orr/add} them: result $= (\text{hi}\!\ll\!16)\,|\,(\text{lo}\!\gg\!16)$.

\begin{lstlisting}
@ Q15 multiply: r0=A (Q15), r1=B (Q15) -> r0=A*B (Q15)
    smull r2, r3, r0, r1   @ r3:r2 = A*B (Q30, 64b)
    lsr   r0, r2, #15      @ shift Q30 -> Q15 (low 32 ok if no overflow)
    bx    lr
\end{lstlisting}

\subsection{HW5 P1 --- Encode 2.75 in UQ3.5 (unsigned, 8b, $n{=}5$)}
\begin{hwp}
$2.75\cdot 2^{5}=2.75\cdot 32 = 88$. Hex: $88=$\textbf{0x58}.
\end{hwp}

\subsection{HW5 P2 --- Decode Q3.4 ($N{=}8$, $n{=}4$)}
\begin{hwp}
\textbf{(a)} 0x77: $U{=}119$, MSB$=$0 $\Rightarrow I{=}119$;
$f=119/16=$\textbf{7.4375}.\\
\textbf{(b)} 0xAA: $U{=}170$, MSB$=$1 $\Rightarrow I=170-256=-86$;
$f=-86/16=$\textbf{$-5.375$}.
\end{hwp}

\subsection{HW5 P3 --- Decode Q2.5 (\texttt{0b1001\_1001})}
\begin{hwp}
$U{=}153$, MSB$=$1 $\Rightarrow I=153-256=-103$;
$f=-103/32=$\textbf{$-3.21875$}.
\end{hwp}

\subsection{HW5 P4 --- Encode in Q3.4 (range $[-8.0,+7.9375]$)}
\begin{hwp}
\textbf{(a)} $0.5\cdot 16=8 \Rightarrow$ \textbf{0x08}.\\
\textbf{(b)} $6.73\cdot 16 = 107.68 \to 108 \Rightarrow$ \textbf{0x6C}
(decode: $108/16=6.75$, err $0.02$).\\
\textbf{(c)} $-2.25\cdot 16=-36$, $+256=220 \Rightarrow$ \textbf{0xDC}.\\
\textbf{(d)} $-4.5\cdot 16=-72$, $+256=184 \Rightarrow$ \textbf{0xB8}.
\end{hwp}

\subsection{HW5 P5 --- Q15 MAC: $f_1f_2+f_3f_4$}
\begin{hwp}
$f_1{=}0.5,\,f_2{=}0.25,\,f_3{=}0.125,\,f_4{=}0.0625$.
\textbf{Encode} ($I{=}f\cdot 2^{15}$): \;
$I_1{=}$0x4000, $I_2{=}$0x2000, $I_3{=}$0x1000, $I_4{=}$0x0800.\\
\textbf{Mult+renorm} ($I_P=(I_AI_B)\!\gg\!15$):
$I_{P1}=2^{14}\!\cdot\!2^{13}\!\gg\!15=2^{12}=$0x1000;
$I_{P2}=2^{12}\!\cdot\!2^{11}\!\gg\!15=2^{8}=$0x0100.\\
\textbf{Add:} $I_A=$0x1100$=4352$.
\textbf{Decode:} $4352/32768=$\textbf{0.1328125} $\checkmark$
($0.5\!\cdot\!0.25+0.125\!\cdot\!0.0625$).
\end{hwp}

\subsection{Worked encode/decode shortcuts}
\begin{ex}
Powers of two: $f=2^{-k}\Rightarrow I=2^{n-k}$ (single-bit code).
Q15 of $0.5=2^{14}=$0x4000; Q15 of $0.0625=2^{11}=$0x0800.
\end{ex}

\begin{ex}
Q15.16 of $3.25$: $I=3.25\cdot 2^{16}=3\cdot 0x10000+0x4000=$\textbf{0x00034000}.
Returned in R0 (32-bit EABI return).
\end{ex}

\subsection{Common pitfalls}
\begin{itemize}
\item Forgetting the $+2^{N}$ wrap on negative encodes (yields wrong sign).
\item Using LSR (logical) instead of ASR for signed renormalize after multiply.
\item Adding numbers of \emph{different} Q-formats without aligning $n$
(shift smaller-$n$ left, or shift larger-$n$ right, before add).
\item Q15$\times$Q15 product is Q30 not Q15 --- must \asm{ASR \#15}.
\item Q15 cannot store $+1.0$ exactly (max $\approx 0.99997$).
\end{itemize}

\subsection{Cookbook table}
\begin{tabular}{@{}lll@{}}
\toprule
op & rule & note \\
\midrule
encode & $I=\mathrm{round}(f\cdot 2^{n})$ & $+2^{N}$ if $I<0$ \\
decode & $f=I\cdot 2^{-n}$ & $I=U-2^{N}$ if MSB \\
add Qm.n & $I_S=I_A+I_B$ & same format only \\
mul Qm.n & $I_P=(I_AI_B)\!\gg\!n$ & ASR for signed \\
Q15 mul & \asm{smull;lsr \#15} & 16$\to$32$\to$16 \\
Q15.16  & 32b, $n{=}16$ & R0 return \\
\bottomrule
\end{tabular}

% ----- 03_hw06_arith_eabi.tex -----
\section{HW6: Arithmetic + EABI Calling}

\subsection{Arithmetic instructions (Cortex-M4)}
\begin{tabular}{@{}lll@{}}
\toprule
Instr & Form & Flags \\
\midrule
\asm{ADD/SUB/RSB} & \asm{Rd,Rn,op2} & none \\
\asm{ADDS/SUBS/RSBS} & \asm{Rd,Rn,op2} & N,Z,C,V \\
\asm{ADC/SBC/RSC} & \asm{Rd,Rn,op2} & uses C in (S sets flags) \\
\asm{MUL/MLA/MLS} & 32$\times$32$\to$lo32 & none (\asm{MULS}: only N,Z) \\
\asm{UMULL/SMULL} & 32$\times$32$\to$64 & none \\
\asm{UMLAL/SMLAL} & RdLo:RdHi += A$\cdot$B & none \\
\asm{UDIV/SDIV} & \asm{Rd,Rn,Rm} & \textbf{never} set flags \\
\bottomrule
\end{tabular}\\
\textbf{RSB:} \asm{rsb Rd,Rn,op2} $=$ op2$-$Rn (reverse). \textbf{MLA} \asm{Rd,Rn,Rm,Ra}: Rd=Ra+Rn$\cdot$Rm. \textbf{MLS} \asm{Rd,Rn,Rm,Ra}: Rd=Ra$-$Rn$\cdot$Rm.\\
\textbf{Flag suffix rule:} append \asm{S} to update NZCV (e.g.\ \asm{ADD}$\to$\asm{ADDS}). Plain \asm{ADD} leaves flags untouched. \asm{UDIV/SDIV} have no S form. Div-by-zero $\to$ Rd=0 (no fault when CFG.DIV\_0\_TRP=0).

\subsection{ARM C / V flag conventions (subtract!)}
\textbf{ADD: } C=1 iff unsigned overflow out of bit 31. V=1 iff signed overflow.\\
\textbf{SUB/CMP: } \emph{C = NOT borrow}. C=1 means \emph{no} borrow (Rn$\ge$op for unsigned); C=0 means borrow occurred. V=1 iff signed overflow.\\
\asm{SBC Rd,Rn,Rm}: Rd $=$ Rn $-$ Rm $-$ \asm{!C} $=$ Rn$-$Rm$-$(1$-$C). So a low-word borrow (C=0) propagates an extra $-1$ into the high word.

\subsection{64-bit ADD / SUB pattern}
\begin{lstlisting}
@ A=R1:R0 (hi:lo), B=R3:R2; result back in R1:R0
mp_add_64bit_s:
    ADDS r0, r2     @ low,  C = unsigned ovf
    ADC  r1, r3     @ high + C
    BX   lr
mp_sub_64bit_s:
    SUBS r0, r2     @ low,  C = !borrow
    SBC  r1, r3     @ high - !C
    BX   lr
\end{lstlisting}
Convention: low word at lower-numbered reg (R0,R2), high at next (R1,R3). The S on the low op is mandatory; the high op must be ADC/SBC, never ADD/SUB.

\subsection{Modulo via UDIV+MUL+SUB (no MOD instr)}
Identity \asm{A\%B = A - (A/B)*B}. Leaf, R0--R2 caller-saved $\Rightarrow$ no PUSH/POP.
\begin{lstlisting}
umod32bit_s:                @ uint32 mod: R0=A,R1=B
    UDIV r2, r0, r1         @ q = A/B
    MUL  r2, r2, r1         @ q*B
    SUB  r0, r0, r2         @ A - q*B
    BX   lr
@ One-instruction form (M4 has MLS):
    UDIV r2, r0, r1
    MLS  r0, r2, r1, r0     @ r0 = r0 - r2*r1
    BX   lr
\end{lstlisting}
For signed modulo use \asm{SDIV} (sign of result follows dividend).

\subsection{EABI / AAPCS register usage}
\begin{tabular}{@{}llp{4.4cm}@{}}
\toprule
Reg & Role & Notes \\
\midrule
R0--R3 & arg / ret, scratch & \emph{caller-saved}; R0(:R1) hold return \\
R4--R11 & local & \emph{callee-saved}; PUSH on entry, POP on exit \\
R12 (IP) & scratch & may be clobbered by veneers \\
R13 (SP) & stack ptr & 8-byte aligned at public boundaries \\
R14 (LR) & link reg & save if you call (BL) anything \\
R15 (PC) & program ctr & POP \{...,pc\} returns \\
\bottomrule
\end{tabular}\\
\textbf{Args:} first 4 word-sized go in R0--R3, rest spill to stack (lowest addr = 5th arg). 64-bit args use an aligned even/odd pair (R0:R1 or R2:R3); cannot split across R3/stack--R3 is left unused if needed for alignment.\\
\textbf{Return:} 32-bit in \textbf{R0}. 64-bit in \textbf{R0:R1} (R0=lo, R1=hi). \texttt{void} returns nothing.\\
\textbf{Leaf rule:} if you don't BL anything and only touch R0--R3,R12, no PUSH/POP needed.

\subsection{Sub-word arg promotion at the CALL SITE}
Caller widens \asm{int8\_t/int16\_t/uint8\_t/uint16\_t} to 32-bit before placing in R0--R3:
\begin{itemize}
\item \textbf{Unsigned} src $\Rightarrow$ \emph{zero-extend} (UXTB/UXTH): upper bits = 0.
\item \textbf{Signed} src $\Rightarrow$ \emph{sign-extend} (SXTB/SXTH): upper bits = MSB of src.
\end{itemize}
Callee assumes the value is already 32-bit clean--reading the low byte without re-extending is wrong if caller failed to extend.

\subsection{Stack \& PUSH/POP rules (full descending)}
SP points at last-pushed word; PUSH pre-decrements, POP post-increments. \textbf{Lowest reg\# stored at lowest address, regardless of brace order.}\\
\asm{PUSH \{r7,r5,r8,r6\}} with SP=0x2000\_1020:\\
\begin{tabular}{@{}ll@{}}
0x2000\_101C & R8 (highest \#) \\
0x2000\_1018 & R7 \\
0x2000\_1014 & R6 \\
0x2000\_1010 & R5 (lowest \#) $\leftarrow$ new SP \\
\end{tabular}\\
SP after $=$ 0x2000\_1020$-$4$\cdot$4 $=$ 0x2000\_1010. POP reverses: lowest addr$\to$lowest reg.\\
\asm{POP \{...,pc\}} replaces the usual final \asm{BX lr}.

\subsection{HW6 P1: PUSH ordering walkthrough}
\begin{hwp}\textbf{P1.1} 4 regs $\cdot$ 4B $=$ 16 $\Rightarrow$ SP$=$0x20001020$-$0x10$=$\textbf{0x20001010}.\\
\textbf{P1.2} R5$=$(1$\ll$5)$+$(1$\ll$10)$=$32$+$1024$=$\textbf{0x0000\_0420}.\\
\textbf{P1.3} Sorted: R5@1010, R6@1014, R7@1018, R8@101C. R6$=$(1$\ll$6)$+$(1$\ll$12)$=$64$+$4096$=$\textbf{0x0000\_1040}.\\
\textbf{P1.4} Little-endian byte@0x20001014 $=$ LSB of 0x1040 $=$ \textbf{0x40}.
\end{hwp}

\subsection{HW6 P2--P3: arg promotion}
\begin{hwp}\textbf{P2 (unsigned, zero-ext):} \cee{fn1((uint8\_t)0xF,(uint8\_t)0xFF,0xFFF,0xFFFF)} into \cee{uint32\_t}\\
R0$=$0x0000\_000F, R1$=$0x0000\_00FF, R2$=$0x0000\_0FFF, R3$=$0x0000\_FFFF.\end{hwp}
\begin{hwp}\textbf{P3 (signed, sign-ext):} \cee{fn2((int8\_t)0xF,(int8\_t)0xFF)} into \cee{int32\_t}\\
0xF: MSB$=$0 $\Rightarrow$ R0$=$\textbf{0x0000\_000F}.\\
0xFF: MSB$=$1 (it's $-1$) $\Rightarrow$ R1$=$\textbf{0xFFFF\_FFFF}. Value $-1$ preserved.\end{hwp}

\subsection{HW6 P4: return-reg + Q15.16 encode}
\begin{hwp}Encode 3.25 in Q15.16: I$=$f$\cdot$2\textsuperscript{16}. $3\!\ll\!16=$0x0003\_0000; $0.25\!\cdot\!65536=$16384$=$0x4000.\\
$I=$\textbf{0x0003\_4000}; 32-bit return $\Rightarrow$ goes in \textbf{R0}.\end{hwp}

\subsection{HW6 P5: 64-bit ADD flag analysis}
\begin{hwp}A$=$R1:R0, B$=$R3:R2.\\
\textbf{T1:} A1$=$0x388$\!\ll\!$24, B1$=$0x289$\!\ll\!$24.\\
lo: 0x88000000$+$0x89000000$=$0x1\_11000000 $\Rightarrow$ R0$=$0x11000000, \textbf{C$=$1}.\\
hi: 3$+$2$+$1$=$6 $\Rightarrow$ R1$=$0x6. Result $=$ \cee{(int64\_t)0x611<<24}.\\
\textbf{T2:} A2$=$0x378$\!\ll\!$24, B2$=$0x279$\!\ll\!$24.\\
lo: 0x78000000$+$0x79000000$=$0xF1000000, fits $\Rightarrow$ \textbf{C$=$0}.\\
hi: 3$+$2$+$0$=$5. Result $=$ \cee{(int64\_t)0x5F1<<24}.\end{hwp}

\subsection{HW6 P6: 64-bit SUB flag analysis}
\begin{hwp}\textbf{T1:} A1$=$0x387$\!\ll\!$28, B1$=$0x28$\!\ll\!$28.\\
lo: 0x70000000$-$0x80000000 $\Rightarrow$ borrow! lo$=$0xF000\_0000, \textbf{C$=$0}.\\
hi: 0x38$-$0x02$-$1$=$0x35. Result $=$ \cee{(int64\_t)0x35F<<28}.\\
\textbf{T2:} A2$=$0x388$\!\ll\!$28, B2$=$0x288$\!\ll\!$28.\\
lo: 0x80000000$-$0x80000000$=$0, no borrow $\Rightarrow$ \textbf{C$=$1}.\\
hi: 0x38$-$0x28$-$0$=$0x10. Result $=$ \cee{(int64\_t)0x100<<28}.\\
\emph{Mantra:} subtract sets C$=$\textbf{!borrow}; SBC subtracts the inverse.\end{hwp}

\subsection{HW6 P7: modulo example}
\begin{hwp}A$=$17, B$=$5: \asm{UDIV} $\to$ q$=$3; \asm{MUL} $\to$ 15; \asm{SUB} $\to$ 17$-$15$=$\textbf{2}. Leaf, no stack frame, R2 is scratch.\end{hwp}

\subsection{Quick gotchas}
\begin{itemize}
\item \asm{MUL} (no S) never sets flags; \asm{MULS} sets only N,Z (C,V undefined). \asm{UMULL/SMULL/UMLAL/SMLAL} never set flags.
\item Reading flags after \asm{UDIV/SDIV} reads stale flags from the previous flag-setting instr.
\item \asm{MLS Rd,Rn,Rm,Ra}: Rd$=$Ra$-$Rn$\cdot$Rm--note the operand order.
\item Signed $\to$ unsigned promotion of a negative byte yields a huge positive uint32 (0xFF\_int8 $\to$ 0xFFFF\_FFFF as int32 $\to$ 4294967295 if reinterpreted unsigned).
\item PUSH brace order is cosmetic; assembler sorts. Don't rely on listed order for memory layout.
\item Full-descending: SP points at \emph{valid} top word; word at [SP] is the most recently pushed.
\end{itemize}

% ----- 04_hw07_shifts_logic.tex -----
\section{HW7: Shifts, Op2, Bitwise Logic}

\subsection{Shift instructions (4 forms + RRX)}
\begin{tabular}{@{}l l l@{}}
\textbf{LSL} & logical shift L & signed/unsigned, 0-fill at LSB \\
\textbf{LSR} & logical shift R & unsigned only, 0-fill at MSB \\
\textbf{ASR} & arith shift R & signed, sign-bit replicated at MSB \\
\textbf{ROR} & rotate right & bits wrap LSB$\to$MSB \\
\textbf{RRX} & rot R thru C & 33-bit rotate by \textbf{1} (C $\to$ MSB, LSB $\to$ C) \\
\end{tabular}

\textbf{S-suffix} (\asm{lsls/lsrs/asrs/rors}): updates flags. \textbf{C = last bit shifted out}; N/Z from result. For \asm{lsls Rd,Rs,\#n} $\Rightarrow$ C = bit$(32{-}n)$ of Rs. For \asm{lsrs Rd,Rs,\#n} $\Rightarrow$ C = bit$(n{-}1)$ of Rs.

\subsection{Operand-2 (flexible second operand)}
Three legal forms in any data-processing instr:
\begin{itemize}
\item \asm{\#imm} -- 8-bit rotated immediate (must encode)
\item \asm{Rm} -- bare register
\item \asm{Rm, \{LSL|LSR|ASR|ROR\} \#n} -- register w/ inline shift (\textbf{free}, no extra cycle)
\end{itemize}
\textbf{Shift count must be immediate} when used as Op2. Standalone \asm{lsl/lsr/asr/ror Rd,Rm,Rs} can take a register count -- only because the shift IS the instruction.

\textbf{GNU quirk:} \asm{add r0, r1, lsl \#3} is illegal even though it looks like \asm{r0 += r0<<3}; you must specify dest+src+Op2 explicitly: \asm{add r0, r0, r1, lsl \#3}. Never omit the Rn slot when Op2 carries a shift.

\subsection{One-instr scaling idioms (decompose $k = 2^n \pm 1$)}
\textbf{Multiply by integer $k$:}
\begin{lstlisting}
lsl r0, r1, #4         @ A = 16*B  (2^n)
add r0, r1, r1, lsl #1 @ A =  3*B
add r0, r1, r1, lsl #2 @ A =  5*B
add r0, r1, r1, lsl #3 @ A =  9*B
add r0, r1, r1, lsl #4 @ A = 17*B  (1+2^n)
rsb r0, r1, r1, lsl #3 @ A =  7*B  (2^n - 1)
rsb r0, r1, r1, lsl #4 @ A = 15*B
\end{lstlisting}
\textbf{Divide / fractional ($k/2^n$, signed):} use ASR.
\begin{lstlisting}
asr r0, r1, #4         @ A = B/16
add r0, r1, r1, asr #4 @ A = 17B/16  = B + B/16
sub r0, r1, r1, asr #4 @ A = 15B/16  = B - B/16
sub r0, r0, r0, asr #2 @ A = 0.75*A  (3/4)
\end{lstlisting}
Recall \asm{rsb Rd,Rn,Op2} $\Rightarrow$ Rd = Op2$-$Rn (reverse). LSR on a signed negative gives garbage; \textbf{always ASR for signed division}.

\subsection{Q31 / Q15.16 renormalization (post-SMULL)}
After \asm{smull Rlo,Rhi,Rn,Rm}, the 64-bit signed product spans $b_{63}\ldots b_0$. Extract Q$m$.$n$ output (bits $b_{m+n+15}\ldots b_n$) with a fused 3-instr renorm:
\begin{lstlisting}
@ Q15.16 result := bits b47..b16 of {Rhi:Rlo}
smull r1, r0, r0, r1   @ Rlo=R1, Rhi=R0
lsr   r0, r1, #16      @ low half: top 16 of LW
orr   r0, r0, r0, lsl #16  @ ... typical pair:
@ canonical:
lsr   r0, r1, #16          @ R0 = (LW >> 16)
orr   r0, r0, r3, lsl #16  @ OR (HW << 16)
\end{lstlisting}
\textbf{Rule:} for Q$m$.$n$, LW $\gg n$, HW $\ll (32{-}n)$, OR. \quad Q31: $\ll 1 / \gg 31$. \quad Q15.16: $\ll 16/\gg 16$. \quad Q16.15: $\ll 17/\gg 15$.

\subsection{Bitwise logic family}
\begin{tabular}{@{}l l l@{}}
\asm{AND Rd,Rn,Op2} & Rd = Rn \& Op2 & mask in \\
\asm{ORR Rd,Rn,Op2} & Rd = Rn $|$ Op2 & set bits \\
\asm{EOR Rd,Rn,Op2} & Rd = Rn $\oplus$ Op2 & toggle bits \\
\asm{BIC Rd,Rn,Op2} & Rd = Rn \& $\sim$Op2 & \textbf{clear} bits set in Op2 \\
\asm{ORN Rd,Rn,Op2} & Rd = Rn $|$ $\sim$Op2 & (no C operator) \\
\asm{MVN Rd,Op2} & Rd = $\sim$Op2 & bitwise NOT \\
\asm{BFC Rd,\#lsb,\#w} & clear $w$ bits at lsb & field clear \\
\asm{BFI Rd,Rn,\#lsb,\#w} & insert low $w$ of Rn at lsb & field insert \\
\end{tabular}

\textbf{C-equivalent gap:} BIC has no direct C operator -- write \cee{a \&= \~{}mask;} (or \cee{a \& \~{}m}). Same for ORN: \cee{a | \~{}b}.

\subsection{APSR access}
\asm{MRS Rd, APSR} -- read flags into Rd. \asm{MSR APSR\_nzcvq, Rs} -- write flags. Used for save/restore around critical sections; rare on exam.

\subsection{Bit-field idioms (memorize)}
\textbf{Replace field of width $w$ at position $s$ with value $v$:}
\begin{lstlisting}
@ a &= ~(MASK << s);   a |= (v << s);
ldr  r1, =((1<<w)-1)   @ width mask (e.g. 0xF for nibble)
bic  r0, r0, r1, lsl #s  @ clear field
ldr  r2, =v
orr  r0, r0, r2, lsl #s  @ insert value
\end{lstlisting}
\textbf{Insert digit at LSB end} (vacated nibble already 0):
\begin{lstlisting}
lsl r0, r0, #4    @ H3H2H1H0 -> H2H1H0 0
orr r0, r0, #3    @ -> H2H1H0 3
\end{lstlisting}
\textbf{Insert digit at MSB end:}
\begin{lstlisting}
lsr r0, r0, #4         @ -> 0 H3H2H1
ldr r1, =0x7000        @ 7000 not 8-bit-rot encodable -> use ldr=
orr r0, r0, r1         @ -> 7 H3H2H1
\end{lstlisting}
\textbf{Set/Clear/Toggle triad:} ORR sets, BIC clears, EOR toggles. MVN inverts.

\subsection{HW7 P1 -- Q15.16 from SMULL}
\begin{hwp}\textbf{P1.} Given \asm{smull r2,r3,r0,r1} produces \{R3:R2\} 64-bit, extract Q15.16 ($b_{47}\ldots b_{16}$):
\begin{lstlisting}
lsr r0, r2, #16        @ low 16 of result = top 16 of LW
orr r0, r0, r3, lsl #16  @ high 16 of result = low 16 of HW
\end{lstlisting}
\end{hwp}

\subsection{HW7 P2 -- LSLS/LSRS \& C flag}
\begin{hwp}\textbf{P2.} R0 = \asm{0xDD} = \asm{0b1101\_1101}.
\begin{lstlisting}
lsls r1, r0, #1   @ R1 = 0x1BA; bit shifted out = R0[31] = 0 -> C=0
lsrs r2, r0, #3   @ R2 = 0x1B;  last bit out = R0[2]  = 1 -> C=1
\end{lstlisting}
After both, C = \textbf{1} (from the last S-instr). Rule: LSLS\#$n$ $\Rightarrow$ C=src$[32{-}n]$; LSRS\#$n$ $\Rightarrow$ C=src$[n{-}1]$.
\end{hwp}

\subsection{HW7 P3 -- Single-instr scaling table}
\begin{hwp}\textbf{P3.} R1=B, write A=R0 in one instruction.
\begin{tabular}{@{}l l@{}}
$A = 16B$  & \asm{lsl r0, r1, \#4} \\
$A = 17B$  & \asm{add r0, r1, r1, lsl \#4} \\
$A = 15B$  & \asm{rsb r0, r1, r1, lsl \#4} \\
$A = B/16$ & \asm{asr r0, r1, \#4} \\
$A = 17B/16$ & \asm{add r0, r1, r1, asr \#4} \\
$A = 15B/16$ & \asm{sub r0, r1, r1, asr \#4} \\
\end{tabular}
\end{hwp}

\subsection{HW7 P4 -- BIC+ORR replace nibble}
\begin{hwp}\textbf{P4.} \cee{uint16\_t a = 0xH3H2H1H0} $\to$ \asm{0xH3H2 A H0} (set bits[7:4]=A).
\begin{lstlisting}
ldr r1, =15
bic r0, r0, r1, lsl #4   @ clear bits 7:4
ldr r2, =10
orr r0, r0, r2, lsl #4   @ OR in 0xA<<4
\end{lstlisting}
C: \cee{(a \& \textasciitilde(0xF<<4)) | (0xA<<4);} or \cee{(a \& 0xFF0F) | 0x00A0;}.
\end{hwp}

\subsection{HW7 P5 -- BIC/ORR/EOR hand-trace}
\begin{hwp}\textbf{P5.} R0=\asm{0xAD}=\asm{1010\_1101}, R1=\asm{0x03}=\asm{0000\_0011}.
\begin{tabular}{@{}l l l@{}}
\asm{lsr r2,r1,\#2} & R2 = 0 & \asm{0x00} \\
\asm{lsl r2,\#2}    & R2 = 0 & \asm{0x00} \\
\asm{bic r3,r0,r1}  & \asm{0xAD \& \~{}0x03 = 0xAD \& 0xFC} & \asm{0xAC} \\
\asm{orr r4,r3,r1}  & \asm{0xAC | 0x03} & \asm{0xAF} \\
\asm{eor r5,r0,r1}  & \asm{0xAD $\oplus$ 0x03} & \asm{0xAE} \\
\end{tabular}
\end{hwp}

\subsection{HW7 P6 -- Digit insert/replace patterns}
\begin{hwp}\textbf{P6(a).} \asm{H3H2H1H0} $\to$ \asm{H3H2 5 H0} (replace H1 with 5):
\begin{lstlisting}
bic r0, r0, #0xF0   @ #0xF0 IS 8-bit-rot encodable
orr r0, r0, #0x50
\end{lstlisting}
\textbf{P6(b).} $\to$ \asm{H2H1H0 3} (shift L, insert 3 at LSB):
\begin{lstlisting}
lsl r0, r0, #4
orr r0, r0, #3
\end{lstlisting}
\textbf{P6(c).} $\to$ \asm{7 H3H2H1} (shift R, insert 7 at MSB):
\begin{lstlisting}
lsr r0, r0, #4
ldr r1, =0x7000      @ 0x7000 not encodable as Op2 imm
orr r0, r0, r1
\end{lstlisting}
\textbf{Why ldr=} for \asm{0x7000}: doesn't fit 8-bit-rotated immediate format. Use \asm{ldr Rx,=const} when an immediate refuses to encode.
\end{hwp}

\subsection{Quick-reference summary}
\begin{itemize}
\item Op2 shift is \textbf{free} -- always prefer \asm{add Rd,Rn,Rm,lsl \#n} over a 2-instr \asm{lsl}+\asm{add}.
\item Always write the \textbf{dest} explicitly when Op2 has a shift -- no implicit \asm{Rd=Rn}.
\item LSL = signed \emph{and} unsigned mul-by-$2^n$. LSR = unsigned div only. ASR = signed div.
\item S-suffix on a shift puts the \textbf{last bit shifted out} into C -- handy single-bit pluck.
\item Field replace = BIC mask $\to$ ORR value-shifted-in. End-insert = shift word, ORR digit.
\item BIC has no C operator: \cee{a \& \~{}m}. ORN: \cee{a | \~{}b}. MVN: \cee{\~{}a}.
\item RRX rotates 33 bits (with C); no count needed -- always 1.
\item APSR via \asm{MRS}/\asm{MSR}; \asm{APSR\_nzcvq} field selector for write.
\end{itemize}

% ----- 05_hw08_nzcv_ccs.tex -----
% =============================================================================
% HW8 --- NZCV flags, condition codes, conditional branches (Lctr 18-19)
% =============================================================================
\section{HW8: NZCV Flags + CCS + Branch}

\subsection{APSR layout \& per-flag meaning}
The four arithmetic flags live in the top nibble of the \textbf{APSR} (Application Program Status Register, view of xPSR):
\begin{center}\scriptsize
\begin{tabular}{@{}cccccl@{}}
\toprule
b31 & b30 & b29 & b28 & b27 & rest \\
\midrule
\textbf{N} & \textbf{Z} & \textbf{C} & \textbf{V} & Q & \cee{0\dots 0} (reserved/IT/GE) \\
\bottomrule
\end{tabular}
\end{center}
\textbf{N} = sign of result = MSB of result word (1 $\Rightarrow$ negative when interpreted signed). \textbf{Z} = (result $==$ 0); pure equality bit, signed/unsigned agnostic. \textbf{C} = unsigned carry-out (add) \emph{or} \textbf{!Borrow} (sub) --- ARM convention is opposite of x86. \textbf{V} = signed overflow: result fell outside $[-2^{N-1},\,2^{N-1}-1]$. Q is the saturation sticky bit (DSP). N\&Z apply to logical ops too; C may be set by shifter; V is unaffected by logical ops.

\subsection{Hand computation in $N$-bit system}
Constants for an $N$-bit ALU: $C_N = 2^N$, $U_{max} = 2^N{-}1$, $S_{min} = -2^{N-1}$, $S_{max} = 2^{N-1}{-}1$. The processor doesn't know signed vs unsigned --- it sets \emph{all four flags every time} and the programmer picks which to consult via the CCS suffix.
\begin{enumerate}
  \item Convert each operand to its $N$-bit unsigned form (negatives: add $C_N$). Same bit pattern for both views.
  \item \textbf{Unsigned table:} compute $u_{tmp}$. If $u_{tmp} > U_{max}$ on add $\Rightarrow$ \textbf{C=1} (carry); subtract $C_N$. If $u_{tmp} < 0$ on sub $\Rightarrow$ \textbf{C=0} (borrow); add $C_N$. Otherwise C matches no-overflow side.
  \item \textbf{Signed table:} compute $s_{tmp}$ from signed operands. If $s_{tmp}$ falls outside $[S_{min},S_{max}]$ $\Rightarrow$ \textbf{V=1}, correct by $\pm C_N$. Else V=0.
  \item \textbf{N} = MSB of corrected binary (same pattern from both tables). \textbf{Z} = 1 iff result is all zeros.
\end{enumerate}
\textbf{ARM C-on-subtract:} after \cee{SUBS}, no borrow $\Rightarrow$ C=1, borrow $\Rightarrow$ C=0. \emph{Mnemonic:} ``C = NOT borrow''. (x86 inverts this --- exam trap if you've used x86.)

\subsection{P1 (42pt) --- 5-bit drills ($C_N{=}32$, $S_{min}{=}{-}16$, $S_{max}{=}15$)}
\begin{center}\scriptsize
\begin{tabular}{@{}llllcccc@{}}
\toprule
op & u-view & s-view & bin & N & Z & C & V \\
\midrule
$(-15){+}({-}5)$ & $17{+}27{=}44{>}31$ & $-20{<}{-}16$ & \cee{0\,1100} & 0 & 0 & \textbf{1} & \textbf{1} \\
$11{-}18$        & $-7{<}0$ borrow     & $11{-}({-}14){=}25$ & \cee{1\,1001} & \textbf{1} & 0 & \textbf{0} & \textbf{1} \\
\bottomrule
\end{tabular}
\end{center}
\begin{hwp}
\textbf{P1a:} both C and V fire because \emph{both} interpretations overflow. Binary \cee{0\,1100} reads as $12$ unsigned and $12$ signed --- they coincidentally agree only because of the wrap. \textbf{P1b:} the operand $18$ is unrepresentable in signed 5-bit, so \cee{10010} reinterprets to $-14$ and the signed subtraction is $11-(-14)=25$, which overflows positive. Sub borrow $\Rightarrow$ C=0 (ARM).
\end{hwp}

\subsection{Flag-only ops: CMP, CMN, TST, TEQ}
These four discard the result and \emph{only} update flags (no Rd):
\begin{center}\scriptsize
\begin{tabular}{@{}lllc@{}}
\toprule
mnem & equiv\ to & op & V affected? \\
\midrule
\cee{CMP Rn,Op2} & \cee{SUBS \_,Rn,Op2}      & subtract & yes \\
\cee{CMN Rn,Op2} & \cee{ADDS \_,Rn,Op2}      & add      & yes \\
\cee{TST Rn,Op2} & \cee{ANDS \_,Rn,Op2}      & bitwise AND & no  \\
\cee{TEQ Rn,Op2} & \cee{EORS \_,Rn,Op2}      & bitwise XOR & no  \\
\bottomrule
\end{tabular}
\end{center}
\textbf{Pseudo:} \asm{cmp r0,\#-5} is illegal as a literal but the assembler rewrites it to \asm{cmn r0,\#5} (subtracting $-5$ = adding $5$). Flags come out identical to the conceptual subtract, so \emph{signed CCS still works correctly}. Use \cee{TST}/\cee{TEQ} for bit-mask checks; they leave V untouched, so V-dependent CCS (\cee{GT}/\cee{LT}/\cee{GE}/\cee{LE}) shouldn't be paired with them.

\subsection{Condition-code suffixes (all 16)}
\begin{center}\scriptsize
\begin{tabular}{@{}lllll@{}}
\toprule
CCS & alias & meaning & flag eqn & domain \\
\midrule
\textbf{EQ} & --   & equal           & Z=1            & both \\
\textbf{NE} & --   & not equal       & Z=0            & both \\
\textbf{CS} & HS   & carry set / $\geq_u$ & C=1       & unsigned \\
\textbf{CC} & LO   & carry clr / $<_u$    & C=0       & unsigned \\
\textbf{MI} & --   & minus / negative& N=1            & rare \\
\textbf{PL} & --   & plus / non-neg  & N=0            & rare \\
\textbf{VS} & --   & ovfl set        & V=1            & rare \\
\textbf{VC} & --   & ovfl clr        & V=0            & rare \\
\textbf{HI} & --   & higher ($>_u$)  & C=1 \& Z=0     & unsigned \\
\textbf{LS} & --   & lower or same ($\leq_u$) & C=0 $\vert$ Z=1 & unsigned \\
\textbf{GE} & --   & $\geq_s$        & N=V            & signed \\
\textbf{LT} & --   & $<_s$           & N$\neq$V       & signed \\
\textbf{GT} & --   & $>_s$           & Z=0 \& N=V     & signed \\
\textbf{LE} & --   & $\leq_s$        & Z=1 $\vert$ N$\neq$V & signed \\
\textbf{AL} & --   & always          & ---            & default \\
\bottomrule
\end{tabular}
\end{center}
\textbf{Mnemonic ``HiLow unSun'':} the H/L family (\cee{HI/HS/LO/LS}) is \emph{unsigned}, the G/L family (\cee{GT/GE/LT/LE}) is \emph{signed}. EQ/NE work either way. Mismatching domain to operand type is the classic exam trap.

\subsection{Match shift type \& CCS to operand domain}
Pair the \emph{shift kind} and the \emph{CCS suffix} to the same domain or you'll get silently wrong answers:
\begin{center}\scriptsize
\begin{tabular}{@{}lll@{}}
\toprule
operand domain & shift for $\div 2^n$ & comparison CCS \\
\midrule
signed   (\cee{int})   & \cee{ASR \#n} (sign-extend) & GT GE LT LE \\
unsigned (\cee{uint})  & \cee{LSR \#n} (zero-fill)   & HI HS LO LS \\
\bottomrule
\end{tabular}
\end{center}
\cee{LSR} on a negative number gives garbage; \cee{ASR} on a large unsigned smears the sign bit. EQ/NE are domain-free.

\subsection{P2 (40pt) --- 6-bit CMP and CCS verify ($C_N{=}64$)}
\textbf{Case (a):} $r0{=}22$, $r1{=}20$. Subtract $\Rightarrow$ result $2$, no borrow, no signed overflow $\Rightarrow$ \textbf{N=0, Z=0, C=1, V=0}. \textbf{Case (b):} $r0{=}{-}20$, $r1{=}{-}22$. Unsigned forms $44$ and $42$; $44{-}42 = 2$ --- identical flags to (a). The ``equal flags from equal magnitudes'' fact is general: $a{-}b$ and $(a{+}C_N){-}(b{+}C_N)$ produce the same result and same NZCV.

\begin{center}\scriptsize
\begin{tabular}{@{}lcccccccc@{}}
\toprule
flags N=0,Z=0,C=1,V=0 & EQ & NE & GT & GE & LT & LE & HI & HS \\
\midrule
evaluate              & F  & T  & T  & T  & F  & F  & T  & T  \\
\bottomrule
\end{tabular}
\end{center}
\begin{hwp}
\textbf{P2 takeaway:} CCS evaluation is \emph{flag-driven, not value-driven}. Two CMPs on totally different operand pairs that happen to yield identical NZCV will branch identically. Therefore the same flag pattern $(0,0,1,0)$ supports \emph{both} ``$22 > 20$'' (signed) and ``$44 >_u 42$'' (unsigned) --- GT and HI both fire. EQ logic is just $Z=1$, NE is $Z=0$ --- domain-independent.
\end{hwp}

\subsection{P3 (20pt) --- max() via negative-logic branch}
C source: \cee{int max(int a,int b)\{ return a{>}{=}b ? a : b; \}}. Negative-logic skeleton: \emph{branch on the opposite CCS to skip the then-block}. For $\geq$ the opposite is $<$, so we use \cee{BLT}.
\begin{lstlisting}
mp_max_ab_i_s:
    cmp   r0, r1            @ a vs b (signed CMP)
    blt   .else             @ NL: a<b skips the then
.then:
    b     .end_if           @ a>=b: r0 already = a, fall to end
.else:
    mov   r0, r1            @ r0 = b
.end_if:
    bx    lr
\end{lstlisting}
\textbf{Why BLT (signed) and not BLO (unsigned)?} \cee{int} operands $\Rightarrow$ signed CCS family. Using \cee{BLO} would silently mis-handle negative $a$ (e.g., $a={-}1$, $b=1$: signed $-1<1$ true, but unsigned $0xFFFFFFFF > 1$ false). Domain mismatch.

\subsection{P4 (30pt) --- 64-bit sum-of-squares with SMLAL}
$\sum_{i=s}^{e} i^2$ overflows 32 bits quickly ($50000^2{=}2.5{\cdot}10^9 > 2^{31}$), so accumulate into a 64-bit pair using the multiply-accumulate variant. \cee{SMLAL Rlo,Rhi,Rn,Rm} computes \cee{\{Rhi:Rlo\} += Rn*Rm} (signed); \cee{UMLAL} is the unsigned twin. EABI returns 64-bit values in \cee{R0} (low) : \cee{R1} (high).
\begin{lstlisting}
mp_range_square_sum_standard_while_s:
    push  {r4, r5}
    ldr   r2, =0            @ sum_lo = 0
    ldr   r3, =0            @ sum_hi = 0
.lp:
    cmp   r0, r1            @ i vs e  (r0 reused as i)
    bgt   .end              @ NL: exit when i>e (signed)
    smlal r2, r3, r0, r0    @ {r3:r2} += i*i  (signed 32x32->64 MAC)
    add   r0, #1
    b     .lp
.end:
    mov   r0, r2            @ lo -> R0
    mov   r1, r3            @ hi -> R1
    pop   {r4, r5}
    bx    lr
\end{lstlisting}
\begin{hwp}
\textbf{Alt:} use \cee{SMULL r4,r5,r0,r0} then \cee{ADDS r2,r2,r4} \cee{ADC r3,r3,r5}; \cee{ADDS}/\cee{ADC} is the generic 64-bit-add idiom (low \cee{ADDS} sets C, high \cee{ADC} adds C). \cee{SMLAL} fuses both into one instruction. Reuse the input \cee{r0} as the loop counter to dodge a callee-saved push.
\end{hwp}

\subsection{Summary / common traps}
\begin{itemize}
  \item Processor sets \emph{all four flags every arithmetic op} --- the programmer chooses C vs V via the CCS suffix.
  \item Sub: \textbf{C = !borrow} (ARM); add: C = carry. Equal magnitudes (e.g. $22{-}20$ vs $({-}20){-}({-}22)$) produce equal flags.
  \item \cee{TST}/\cee{TEQ} \emph{don't touch V} --- never follow them with V-dependent CCS (\cee{GT}/\cee{LT}/\cee{GE}/\cee{LE}).
  \item \cee{cmp r0,\#-5} pseudo-assembles as \cee{cmn r0,\#5}; signed CCS still semantically valid.
  \item Match operand domain to both shift kind (\cee{ASR} signed / \cee{LSR} unsigned) and CCS family (G/L signed / H/L unsigned).
  \item Negative-logic branch reads cleaner: \cee{cmp;\,b\textit{!cond}\,.else} skips the then-block.
  \item For 64-bit accumulate use \cee{SMLAL/UMLAL} (or \cee{ADDS}/\cee{ADC} pair); return low in R0, high in R1.
\end{itemize}

% ----- 06_hw09_ldr_str.tex -----
\section{HW9: LDR/STR Addressing Modes}

\subsection{Three immediate-offset forms}
Let \asm{Ra}$=$base, \asm{\#os}$=$signed small immediate.
\begin{tabular}{@{}lll@{}}
\toprule
syntax & EA & R\textsubscript{a} after \\
\midrule
\asm{[Ra,\#os]}    & \asm{Ra+os} & unchanged (basic) \\
\asm{[Ra,\#os]!}   & \asm{Ra+os} & $\leftarrow$\asm{Ra+os} (pre) \\
\asm{[Ra],\#os}    & \asm{Ra}    & $\leftarrow$\asm{Ra+os} (post) \\
\bottomrule
\end{tabular}
\textbf{Mnem:} \asm{!}$=$update first then load ($\equiv$\cee{*++p});
trailing \asm{,\#os}$=$load first then update ($\equiv$\cee{*p++});
neither $=$\cee{*(p+k)}.

\subsection{Load-size suffixes}
\begin{tabular}{@{}llll@{}}
\toprule
op & bits & extend & C type \\
\midrule
\asm{LDR}   & 32 & ---           & \cee{int32\_t} \\
\asm{LDRH}  & 16 & zero          & \cee{uint16\_t} \\
\asm{LDRSH} & 16 & sign          & \cee{int16\_t}  \\
\asm{LDRB}  & 8  & zero          & \cee{uint8\_t}  \\
\asm{LDRSB} & 8  & sign          & \cee{int8\_t}   \\
\bottomrule
\end{tabular}
\textbf{STR} has \emph{no} sign variants: \asm{STR/STRH/STRB} only --- write
just truncates the low bits of the source register.

\subsection{Sign-extension trap}
\begin{tabular}{@{}ll@{}}
\toprule
load mem & R\textsubscript{d} (32b) \\
\midrule
LDRB 0x80      & 0x00000080 \\
LDRSB 0x80     & 0xFFFFFF80 \\
LDRSB 0x40     & 0x00000040 \\
LDRH 0xB0A0    & 0x0000B0A0 \\
LDRSH 0xB0A0   & 0xFFFFB0A0 \\
\bottomrule
\end{tabular}
Sign bit$=$bit-7 (B) or bit-15 (H). If set under SB/SH, top 24/16
bits become \asm{1}.

\subsection{Register-offset (scaled index)}
\asm{[Ra, Ri, LSL \#n]} $\Rightarrow$ EA $=$ \asm{Ra + (Ri$\!\ll\!$n)}.
$n=\log_2(\text{elem size})$:
\asm{\#0}/B,SB; \asm{\#1}/H,SH; \asm{\#2}/word; \asm{\#3}/dword.
\textbf{No writeback form} --- \asm{Ra} never updates.
This is the \cee{p[i]} idiom.

\subsection{LDRD / STRD --- 64-bit pair}
\asm{LDRD Rt,Rt2,[Ra,...]} loads two consecutive words into
\textbf{consecutive regs} (Rt, Rt2$=$Rt+1).
Forms: basic \asm{[Ra,\#os]}, pre \asm{[Ra,\#os]!}, post \asm{[Ra],\#os}.
\textbf{NO reg-offset form.}
Rt$=$word@EA, Rt2$=$word@EA$+4$.

\subsection{C $\leftrightarrow$ ASM cookbook}
\begin{tabular}{@{}ll@{}}
\toprule
C idiom & ARM form \\
\midrule
\cee{*p}                  & \asm{LDR r,[Ra]}            \\
\cee{*(p+k)}              & \asm{LDR r,[Ra,\#k*sz]}     \\
\cee{p[i]}                & \asm{LDR r,[Ra,Ri,LSL \#n]} \\
\cee{*p++}                & \asm{LDR r,[Ra],\#sz}       \\
\cee{*++p}                & \asm{LDR r,[Ra,\#sz]!}      \\
\cee{*p = v} (no upd)     & \asm{STR r,[Ra]}            \\
\cee{*p++ = v}            & \asm{STR r,[Ra],\#sz}       \\
\cee{*++p = v}            & \asm{STR r,[Ra,\#sz]!}      \\
\bottomrule
\end{tabular}
\textbf{What's modified:}\;
prefix \asm{++} on \cee{*p} $\Rightarrow$ value in memory changes
$\Rightarrow$ need explicit STR writeback;
postfix on \emph{pointer} $\Rightarrow$ pointer changes (post-index);
both ($\cee{*p8b++ = ++*p8a++}$) $\Rightarrow$ LDRB post + ADD +
STRB \asm{[Ra,\#-1]} writeback.

\subsection{HW9 P1 --- Trace through 16 bytes}
Memory: 16 bytes \asm{0x00..0x0F} at \asm{0x20000000} (LE words:
\asm{ipt+0=0x03020100}, \asm{+4=0x07060504},
\asm{+8=0x0B0A0908}, \asm{+12=0x0F0E0D0C}).
\begin{lstlisting}
ldr r4, =0x20000000     @ (1) R4=0x20000000
ldr r0, [r4, #4]        @ (2) basic
ldr r1, [r4, #2]!       @ (3) pre-idx
ldr r2, [r4], #4        @ (4) post-idx
ldr r3, [r4]            @ (5) basic, no offset
\end{lstlisting}
\begin{tabular}{@{}llll@{}}
\toprule
step & EA & dest & R4 after \\
\midrule
(1) & ---        & ---           & 0x20000000 \\
(2) & 0x20000004 & R0=0x07060504 & 0x20000000 \\
(3) & 0x20000002 & R1=0x05040302 & 0x20000002 \\
(4) & 0x20000002 & R2=0x05040302 & 0x20000006 \\
(5) & 0x20000006 & R3=0x09080706 & 0x20000006 \\
\bottomrule
\end{tabular}
\textbf{Key:} pre-idx (3) loads from \emph{new} R4; post-idx (4) loads
from \emph{old} R4. Loads at non-word-aligned addresses (0x...02) work
fine on Cortex-M but cross word boundaries.

\subsection{HW9 P2 --- ASM $\to$ C prototypes}
\begin{lstlisting}
mp_task1:                  mp_task2:
  ldrsh r2,[r0],#4           ldrh  r2,[r0,#2]!
  str   r2,[r1,#0]           str   r2,[r1,#0]
  ldrsh r3,[r0],#4           ldrh  r3,[r0,#2]!
  str   r3,[r1,#4]           str   r3,[r1,#4]
  bx    lr                   bx    lr
\end{lstlisting}
\textbf{Task1:} \cee{void mp\_task1(int16\_t *pSrc, int32\_t *pDst)} ---
LDRSH$\Rightarrow$signed src; STR (word)$\Rightarrow$word dst.
Post-idx \asm{\#4} skips \emph{every other halfword} (stride$=2$ in
\cee{int16\_t} units).
\textbf{C:} \cee{*p32 = (int32\_t)*p16s; p16s += 2;} \newline \cee{
*(p32+1) = (int32\_t)*p16s; p16s += 2;}

\textbf{Task2:} \cee{void mp\_task2(uint16\_t *pSrc, int32\_t *pDst)} ---
LDRH$\Rightarrow$unsigned. Pre-idx \asm{\#2} advances first $\Rightarrow$
\emph{skips arr[0]}, reads \cee{arr[1]} then \cee{arr[2]}.
\textbf{C:} \cee{p16u += 1; *p32 = (int32\_t)*p16u;} \newline \cee{
p16u += 1; *(p32+1) = (int32\_t)*p16u;}

\subsection{HW9 P3 --- Four tasks, same data}
\cee{ipt[16]} initialized \cee{ipt[i]=i<<4} ($0x00,0x10,\ldots,0xF0$).
Words (LE, treating ipt as \cee{int32\_t*}):
\begin{tabular}{@{}ll@{}}
\toprule
addr & word \\
\midrule
\cee{ipt+0}  & 0x30201000 \\
\cee{ipt+4}  & 0x70605040 \\
\cee{ipt+8}  & 0xB0A09080 \\
\cee{ipt+12} & 0xF0E0D0C0 \\
\bottomrule
\end{tabular}
Entry: \asm{R0}$=$\asm{0x20001010} (\cee{ipt}),
\asm{R1}$=$\asm{0x20001040} (\cee{opt}).
T4 also gets \asm{R2}$=1$.

\subsubsection*{T1 --- basic offset (R0 unchanged)}
\begin{lstlisting}
ldr r2,[r0,#4]      @ R2=0x70605040
str r2,[r1,#0]      @ opt[0]
ldr r2,[r0,#0xC]    @ R2=0xF0E0D0C0
str r2,[r1,#4]      @ opt[1]
\end{lstlisting}
\textbf{Print:} \asm{Out1=0x70605040, Out2=0xF0E0D0C0}.\;
\textbf{After:} R0$=$0x20001010, R2$=$0xF0E0D0C0.\;
\textbf{C:} \cee{int32\_t *p=(int32\_t*)ipt; opt[0]=p[1]; opt[1]=p[3];}

\subsubsection*{T2 --- pre-index ($\times 2$)}
\begin{lstlisting}
ldr r2,[r0,#4]!     @ R0=0x...14, R2=0x70605040
str r2,[r1,#0]
ldr r2,[r0,#4]!     @ R0=0x...18, R2=0xB0A09080
str r2,[r1,#4]
\end{lstlisting}
\textbf{Print:} \asm{Out1=0x70605040, Out2=0xB0A09080}.\;
\textbf{After:} R0$=$0x20001018, R1$=$0x20001040 (unchanged).\;
\textbf{C:} \cee{p++; opt[0]=*p; p++; opt[1]=*p;}

\subsubsection*{T3 --- post-index ($\times 2$)}
\begin{lstlisting}
ldr r2,[r0],#4      @ R2=0x30201000, R0=0x...14
str r2,[r1,#0]
ldr r2,[r0],#4      @ R2=0x70605040, R0=0x...18
str r2,[r1,#4]
\end{lstlisting}
\textbf{Print:} \asm{Out1=0x30201000, Out2=0x70605040}.\;
\textbf{After:} R0$=$0x20001018.\;
\textbf{C:} \cee{opt[0]=*p; p++; opt[1]=*p; p++;}\\
\textbf{T2 vs T3:} pre starts at \cee{ipt+4}, post starts at \cee{ipt+0}
--- pre and post end at \emph{same} R0 but read \emph{different} words.

\subsubsection*{T4 --- register offset \asm{LSL \#2}}
Entry R2$=1$.
\begin{lstlisting}
ldr r3,[r0,r2,lsl #2]   @ ipt+4, R3=0x70605040
str r3,[r1,#0]
add r2,#1               @ R2=2
ldr r3,[r0,r2,lsl #2]   @ ipt+8, R3=0xB0A09080
str r3,[r1,#4]
add r2,#1               @ R2=3
\end{lstlisting}
\textbf{Print:} \asm{Out1=0x70605040, Out2=0xB0A09080}.\;
\textbf{After:} R0$=$0x20001010 (unchanged!), R2$=$3, R3$=$0xB0A09080.\;
\textbf{C:} \cee{int i=1; opt[0]=p[i]; i++; opt[1]=p[i]; i++;}

\subsection{Pre vs post --- ending pointer is the same}
For two consecutive same-stride loads from base \asm{Ra}:
\textbf{pre} reads \asm{[Ra+s, Ra+2s]} ending at \asm{Ra+2s};
\textbf{post} reads \asm{[Ra, Ra+s]} ending at \asm{Ra+2s}.
Same \emph{final} pointer, different \emph{data}.

\subsection{Common pitfalls}
\begin{itemize}\setlength{\itemsep}{0pt}
\item Pre \asm{[Ra,\#os]!} vs post \asm{[Ra],\#os}: \asm{!} \emph{inside}
      brackets $=$ EA changed inside the access.
\item Stride matches elem size: \cee{int16\_t*}+1$=$\asm{\#2};
      \cee{int32\_t*}+1$=$\asm{\#4}.
\item LDRSB/LDRSH on 0x80/0x8000 $\to$ 0xFFFFFF80/0xFFFF8000.
\item Reg-offset has no writeback; to advance use
      \asm{ADD Ra,Ra,Ri,LSL \#n}.
\item LDRD/STRD: consecutive regs, no reg-offset form.
\item ``Follow assembly's intention'': if .s treats \cee{uint8\_t*} as
      \cee{int32\_t*}, stride is 4 bytes regardless of declared type.
\item \cee{*p8b++ = ++*p8a++} $\Rightarrow$ \asm{LDRB+ADD+STRB[r0,\#-1]}
      writeback to original address.
\end{itemize}

\subsection{Decision table}
\begin{tabular}{@{}lll@{}}
\toprule
need & form & note \\
\midrule
read element, keep ptr  & \asm{[Ra,\#k]}      & basic    \\
walk forward (post)     & \asm{[Ra],\#sz}     & \cee{*p++}    \\
skip first, then walk   & \asm{[Ra,\#sz]!}    & \cee{*++p}    \\
random index            & \asm{[Ra,Ri,LSL \#n]}& \cee{p[i]}; no upd \\
two words at once       & \asm{LDRD r,r+1,...}& consec. reg pair \\
signed byte/half        & \asm{LDRSB/LDRSH}   & sign-ext to 32 \\
unsigned byte/half      & \asm{LDRB/LDRH}     & zero-ext to 32 \\
store (any size)        & \asm{STR/STRH/STRB} & no sign variants \\
\bottomrule
\end{tabular}

% ----- 07_hw10_ptr_fn_call.tex -----
\section{HW10: C Pointers + LDR/STR + Function Calls}

\subsection{C precedence (rows that matter)}
\begin{tabular}{@{}clll@{}}
\toprule
L & ops & assoc & note \\
\midrule
1 & \cee{() [] . -> ++ -- (post)} & L$\to$R & postfix inc/dec \\
2 & \cee{++ -- (pre)}, \cee{* \& (cast)} & R$\to$L & deref, addr-of \\
14 & \cee{= += ...} & R$\to$L & assignment \\
\bottomrule
\end{tabular}
\textbf{Rule:} postfix \cee{p++} (L1) yields OLD pointer; deref/prefix-\cee{++}/cast share L2 R$\to$L. Decode right-to-left.

\subsection{Idiom $\to$ ASM cookbook}
\begin{tabular}{@{}lll@{}}
\toprule
C & ASM & note \\
\midrule
\cee{v=*(p+k)}  & \asm{ldrb r0,[r1,\#k]}   & basic, p unchanged \\
\cee{v=*++p}    & \asm{ldrb r0,[r1,\#1]!}  & pre (p first) \\
\cee{v=*p++}    & \asm{ldrb r0,[r1],\#1}   & post (load first) \\
\cee{*(p+k)=v}  & \asm{strb r0,[r1,\#k]}   & basic store \\
\cee{*p++=v}    & \asm{strb r0,[r1],\#1}   & post store \\
\cee{++(*p)}    & ldr;add\#1;str           & READ-MOD-WRITE \\
\cee{(*p)++}    & ldr;mov;add\#1;str       & return old \\
\cee{++*p++}    & ldr post;add;str\asm{[\#-1]} & WB to OLD \\
\cee{*p++=*q++} & ldr post; str post       & two ptrs walk \\
\bottomrule
\end{tabular}
\textbf{Hook:} \texttt{!}$=$urgent update BEFORE; trailing \texttt{\#}$=$update AFTER.

\subsection{Type suffix table (load only --- stores zero/sign agnostic)}
\begin{tabular}{@{}llll@{}}
\toprule
sfx & C type & ext & sample \\
\midrule
(none) & \cee{int32\_t} & --- & \asm{ldr} \\
B  & \cee{uint8\_t}  & zero $\to$ 32 & 0x80$\to$0x0000\_0080 \\
SB & \cee{int8\_t}   & sign $\to$ 32 & 0x80$\to$0xFFFF\_FF80 \\
H  & \cee{uint16\_t} & zero $\to$ 32 & 0xB0A0$\to$0x0000\_B0A0 \\
SH & \cee{int16\_t}  & sign $\to$ 32 & 0xB0A0$\to$0xFFFF\_B0A0 \\
\bottomrule
\end{tabular}
Sign-extension test: MSB of the loaded chunk (bit 7 for SB, bit 15 for SH).
\textbf{STR has no S/U variant} (you store the bits you have).

\subsection{Pointer arithmetic scaling}
Increment of typed pointer scales by \cee{sizeof(*p)}: \cee{p8++}$=$+1B,
\cee{p16++}$=$+2B, \cee{p32++}$=$+4B. Register-offset shift exponent matches
log$_2$(elem): \asm{[Rb,Ri,LSL \#0/1/2/3]} for byte/hw/word/dword.
\textbf{Typed difference} \cee{q-p} returns \emph{elements}, not bytes
(divide-by-size baked in).

\subsection{HW10 P1 --- four-line memory \& pointer trace}
Array: \cee{uint8\_t my\_arr[16]} at \texttt{0x2000\_0100}, \cee{my\_arr[i]=2i}.
\cee{p8a=0x...100}, \cee{p8b=0x...10C}.
\begin{tabular}{@{}lllll@{}}
\toprule
ln & RHS & memory effect & p8a$\to$ & p8b$\to$ \\
\midrule
4 & \cee{*p8b++=*(p8a+2)}     & rd 0x102$=$0x04; \texttt{0x10C}$\leftarrow$0x04 & 0x100 & 0x10D \\
5 & \cee{*p8b++=++*(p8a+4)}   & 0x104:0x08$\to$0x09; \texttt{0x10D}$\leftarrow$0x09 & 0x100 & 0x10E \\
6 & \cee{*p8b++=++*p8a++}     & 0x100:0x00$\to$0x01; \texttt{0x10E}$\leftarrow$0x01 & 0x101 & 0x10F \\
7 & \cee{*p8b++=*p8a++}       & rd 0x101$=$0x02; \texttt{0x10F}$\leftarrow$0x02 & 0x102 & 0x110 \\
\bottomrule
\end{tabular}

\textbf{Bytes at 0x10C--0x10F:} \texttt{0x04,0x09,0x01,0x02}.
\textbf{Mem changes by RHS:} L4 none, L5 0x104(0x08$\to$0x09), L6 0x100(0x00$\to$0x01), L7 none.

\textbf{Line 6 dissection (\cee{++*p8a++}, the trickiest):}
postfix \cee{p8a++} (L1) gives OLD pointer 0x100; deref reads 0x00; prefix
\cee{++} adds 1, writes 0x01 BACK to 0x100. So memory at 0x100 changes; the
loaded value is 0x01; p8a now 0x101.

\subsection{HW10 P1.3 --- ASM (leaf, \cee{my\_fn(uint8\_t*, uint8\_t*)})}
\begin{lstlisting}
my_fn:                       @ R0=p8a, R1=p8b (caller-saved)
  @ L4: *p8b++ = *(p8a+2);
    ldrb r2, [r0, #2]          @ basic offset, p8a stays
    strb r2, [r1], #1          @ post-idx: *p8b=r2; p8b++
  @ L5: *p8b++ = ++*(p8a+4);
    ldrb r2, [r0, #4]
    add  r2, r2, #1            @ prefix-inc the value
    strb r2, [r0, #4]          @ WRITEBACK -- don't forget!
    strb r2, [r1], #1
  @ L6: *p8b++ = ++*p8a++;
    ldrb r2, [r0], #1          @ post-idx: r2=*OLD p8a; R0+=1
    add  r2, r2, #1
    strb r2, [r0, #-1]         @ writeback to OLD addr (R0-1)
    strb r2, [r1], #1
  @ L7: *p8b++ = *p8a++;
    ldrb r2, [r0], #1
    strb r2, [r1], #1
    bx   lr                     @ leaf: no PUSH/POP, no LR save
\end{lstlisting}

\textbf{Three patterns to memorize:}
(1) \cee{++(*(p+k))} $=$ ldrb / add\#1 / strb-to-same-offset / strb-to-dest;
(2) \cee{++*p++}     $=$ ldrb post-idx / add / strb \asm{[\#-1]} / strb post-idx;
(3) \cee{*p++}       $=$ ldrb post-idx alone (no writeback).

\subsection{AAPCS: who-saves-what (the only table you need)}
\begin{tabular}{@{}lll@{}}
\toprule
reg & role & preserved? \\
\midrule
R0--R3 & args + return value & no (caller-saved, scratch) \\
R4--R11& callee scratch       & \textbf{yes} (PUSH if used) \\
R12 (IP)& scratch              & no \\
R13 (SP)& stack pointer        & yes (8-byte aligned at boundary) \\
R14 (LR)& return addr          & save if you BL (stem) \\
R15 (PC)& program counter      & --- \\
\bottomrule
\end{tabular}

\textbf{Caller widens narrow types} (\cee{int8\_t/int16\_t}) to 32b before
placing in R0--R3: signed args sign-ext, unsigned zero-ext. 64-bit args use
register pair (lo, hi), even-aligned (R0:R1 or R2:R3).

\subsection{Leaf vs Stem function templates}
\textbf{Leaf} (no BL inside, $\le$4 scratch regs needed): no push, no pop.
\begin{lstlisting}
leaf_fn:                  @ R0..R3 in, R0 out
    @ work in R0--R3, R12
    bx   lr
\end{lstlisting}

\textbf{Stem} (calls others or needs R4--R11): push LR + any callee-saved used.
Stash-args pattern: copy R0--R3 into R4--R7 \emph{before} BL clobbers them.
\begin{lstlisting}
stem_fn:
    push {r4-r7, lr}        @ save callee + return addr
    mov  r4, r0             @ stash args (R0..R3 die on next BL)
    mov  r5, r1
    mov  r6, r2
    mov  r7, r3
    @ ... bl other_fn; arg in R0; result in R0 ...
    @ recover args from R4--R7 between calls
    pop  {r4-r7, pc}        @ pop PC = return; replaces BX LR
\end{lstlisting}
Push/pop list \emph{must match}; pushing odd \# of regs $\to$ pad for 8B align
(usually push LR alongside even count). \asm{POP \{...,pc\}} returns directly.

\subsection{HW10 P2 --- four-task LDRD/SB/SH/H trace}
Setup: \cee{ipt[i]=i<<4} $\to$ bytes 0x00,0x10,0x20,...,0xF0. Word at ipt+0
$=$ 0x30201000 (little-endian: LSB at lowest addr). Halfwords at
ipt+0/+2/.../+14 $=$ 0x1000,0x3020,0x5040,0x7060,0x9080,0xB0A0,0xD0C0,0xF0E0.
Entry R0$=$0x2000\_1010 (ipt), R1$=$0x2000\_1040 (opt).

\textbf{T1: LDRD pre-idx (8B$\to$2 regs).}
\asm{ldrd r2,r3,[r0,\#8]!} : R0$+$=8 first, then load lo word at \texttt{[R0]}
$=$ 0xB0A09080, hi word at \texttt{[R0+4]} $=$ 0xF0E0D0C0.
\asm{strd r2,r3,[r1,\#0]} writes opt[0],opt[1].
\textbf{Print:} \texttt{Out1=0xB0A09080, Out2=0xF0E0D0C0}.

\textbf{T2: LDRSB pre-idx, sign-extend.}
\asm{ldrsb r2,[r0,\#4]!}: R0$\to$ipt+4, byte 0x40 (bit7$=$0) $\to$ R2$=$0x0000\_0040.
\asm{ldrsb r3,[r0,\#4]!}: R0$\to$ipt+8, byte 0x80 (bit7$=$1) $\to$ R3$=$0xFFFF\_FF80.
After: R0$=$\textbf{0x2000\_1018}, R1$=$0x2000\_1040 (STRD has no \texttt{!}).
\textbf{Print:} \texttt{Out1=0x40, Out2=0xFFFFFF80}.

\textbf{T3: LDRSH post-idx, odd stride 10.}
\asm{ldrsh r2,[r0],\#10}: load hw at ipt+0$=$0x1000 (bit15$=$0) $\to$
R2$=$0x0000\_1000; THEN R0$+$=10.
\asm{ldrsh r3,[r0],\#4}: hw at ipt+10$=$0xB0A0 (bit15$=$1) $\to$
R3$=$0xFFFF\_B0A0; R0$+$=4.
After: R0$=$\textbf{0x2000\_101E} (unaligned!).
\textbf{Print:} \texttt{Out1=0x1000, Out2=0xFFFFB0A0}.

\textbf{T4: LDRH register-offset, zero-ext.} Called \cee{task4(ipt,opt,1)}, R2$=$1.
\asm{ldrh r3,[r0,r2,LSL \#1]}: EA$=$ipt$+$2, hw 0x3020 $\to$ R3$=$0x0000\_3020;
\asm{str r3,[r1,\#0]}; \asm{add r2,\#1} (R2$=$2); reload at ipt$+$4 $=$
0x5040 $\to$ R3$=$0x0000\_5040; \asm{str r3,[r1,\#4]}; \asm{add r2,\#1} (R2$=$3).
\textbf{Print:} \texttt{Out1=0x3020, Out2=0x5040}. \textbf{R2$=$0x3, R3$=$0x5040.}

\subsection{Sign-extend pivots: LOAD vs CALL}
\textbf{LOAD:} LDRSB/LDRSH test bit 7/15 of loaded data, widen to 32b. LDRB/LDRH zero-pad.
\textbf{CALL:} caller widens narrow C arg per type signedness BEFORE \asm{BL}. Two independent extensions.

\subsection{HW10 P2 --- ASM$\to$C reverse}
\textbf{T1} 8B copy: \cee{*((int64\_t*)opt)=*((int64\_t*)ipt+1)} (or word-by-word with \cee{(int32\_t*)}).
\textbf{T2} pre-inc \cee{int8\_t*}: \cee{p8+=4; opt[0]=*p8; p8+=4; opt[1]=*p8;}.
\textbf{T3} post-inc \cee{int16\_t*}: \cee{opt[0]=*p16; opt[1]=*(p16+5);} (\#10$=$5 elem).
\textbf{T4} idx \cee{uint16\_t*}: \cee{int i=1; opt[0]=*(p16u+i); i++; opt[1]=*(p16u+i); i++;}.

\subsection{Array-loop $+$ call pattern (compute index $\to$ ldr/bl/str)}
Idiom: \cee{for(i;..) arr[i]=fn(arr[i],...)} compiles to:
\begin{lstlisting}
    @ R5=&arr (base, callee-saved), R8=i (callee-saved)
loop:
    ldr  r0, [r5, r8, lsl #2]   @ R0 = arr[i]   (32b elements)
    bl   fn                      @ R0 = result; clobbers R0--R3,R12
    str  r0, [r5, r8, lsl #2]   @ arr[i] = R0
    add  r8, r8, #1
    cmp  r8, r9                  @ R9 = N (callee-saved)
    blt  loop
\end{lstlisting}
\textbf{Why callee-saved (R4--R11) for base/index/N:} they survive across
\asm{BL fn}. Putting them in R0--R3/R12 would corrupt on every call.

\subsection{Fast-failure modes}
\begin{itemize}
\item Forgetting STRB writeback after \cee{++(*p)}: needs \asm{ldr;add;str} then dest store.
\item \asm{[r0,\#4]!} (pre, urgent BEFORE) vs \asm{[r0],\#4} (post, AFTER).
\item LDRB on \cee{int8\_t} $\to$ wrong zero-pad; use LDRSB.
\item STRD/LDRD w/o \texttt{!} doesn't update Ra (T1 R1 unchanged; T2 R0 advances).
\item Reg-offset shift: LSL \#0/1/2/3 for byte/hw/word/dword.
\item No \asm{push \{lr\}} in stem $\to$ \asm{BL} clobbers LR, return broken.
\end{itemize}

\subsection{One-line summary table}
\begin{tabular}{@{}lll@{}}
\toprule
goal & instruction & C \\
\midrule
\cee{*(p+k)}      & \asm{ldr\{B/H\}\{S\} Rd,[Ra,\#os]} & basic, p stays \\
\cee{*++p}        & \asm{...,[Ra,\#os]!}                & pre, p$+$=os first \\
\cee{*p++}        & \asm{...,[Ra],\#os}                 & post, load then p$+$=os \\
\cee{*(p+i)} typed& \asm{...,[Ra,Ri,LSL \#k]}           & k$=$log$_2$elem \\
8B copy           & \asm{ldrd}/\asm{strd}               & 64b, two regs \\
sign-ext load     & \asm{ldrsb}/\asm{ldrsh}             & test bit 7/15 \\
leaf return       & \asm{bx lr}                          & no push \\
stem return       & \asm{pop \{...,pc\}}                & pushed LR \\
\bottomrule
\end{tabular}

% ----- 08_hw11_if_flow.tex -----
\section{HW11: If-Based Flow Control (PL/NL/CEX)}

\subsection{Three styles for \cee{if-else} in ASM}
Given C \cee{if (cond) then; else else;}, three idiomatic ASM templates (Lctr 25):
\begin{tabular}{@{}l l l@{}}
\toprule
Style & Branch on & Layout \\
\midrule
\textbf{PL} (positive logic) & \textbf{same} CCS as C & else first, fall thru / B end / then \\
\textbf{NL} (negative logic) & \textbf{inverse} CCS  & then first, fall thru / B end / else \\
\textbf{CEX} (cond exec)     & no branch (IT block)  & all instrs in straight line \\
\bottomrule
\end{tabular}\\
NL reads closest to C (then-block falls through, matches \cee{if(c){...}}). PL is "branch when C-true". CEX is best when each arm is $\le$ 4 instructions total. CCS in \asm{IT..} = NL cond (the inverse of the C test).

\subsection{IT block syntax + hard rules}
\asm{IT\{x\{y\{z\}\}\} cond} -- up to \textbf{4} conditional instrs. First slot always \asm{T} (matches \asm{cond}); each later slot is \asm{T} (same cond) or \asm{E} (\textbf{inverse} cond).
\textbf{Common forms (T = same cond, E = inverse):} \asm{IT} (T), \asm{ITT} (T,T), \asm{ITE} (T,E), \asm{ITTT} (T,T,T), \asm{ITTE} (T,T,E), \asm{ITEE} (T,E,E), \asm{ITTEE} (T,T,E,E), \asm{ITTTT} (T,T,T,T).\\
\textbf{Hard rules (Lctr 24, p.\ 9):} max \textbf{4} cond instrs; \textbf{cond B} (\asm{Bcc}) only in \textbf{LAST} slot; \textbf{CBZ/CBNZ FORBIDDEN} inside IT (use \asm{cmp+Bcc}); each slot's CCS must match T/E role; cond LDR/STR (pre/post-idx) \textbf{legal} in IT.

\subsection{CBZ / CBNZ}
\asm{CBZ Rn,lbl} = "if Rn==0 goto lbl"; \asm{CBNZ Rn,lbl} = "if Rn!=0". \textbf{Forward only}, max \textbf{126 B} ahead. \asm{Rn} restricted to \textbf{R0--R7}. \textbf{Does NOT set flags}. Replaces \asm{cmp Rn,\#0 ; beq lbl} (1 instr vs 2). Forbidden inside IT. Use for null/loop guards.

\subsection{Sign-aware CCS (HW 11 always signed)}
Signed: \asm{GT/GE/LE/LT}; unsigned: \asm{HI/HS/LS/LO}. \textbf{Inverse pairs:} GT$\leftrightarrow$LE, GE$\leftrightarrow$LT, EQ$\leftrightarrow$NE. NL branches on the inverse. \textbf{Type drives 3 things:} CCS (GT/LE vs HI/LS), shift (ASR vs LSR), load (SB/SH vs B/H).

\subsection{Compound conditions}
\textbf{OR} (\cee{a||b}) -- short-circuit with TWO \asm{CMP}s, both branch to the SAME \asm{then} label; fall-through = else:
\begin{lstlisting}
    cmp r0, #20
    ble then         @ a true -> then
    cmp r0, #25
    bge then         @ b true -> then
    b   else_        @ neither -> else
\end{lstlisting}
\textbf{AND} (\cee{a\&\&b}) -- via De Morgan, branch to ELSE on inverse of either:
\begin{lstlisting}
    cmp r0, #LO
    blt else_        @ !a -> else (skip then)
    cmp r0, #HI
    bgt else_        @ !b -> else
    @ both true: fall through to then
\end{lstlisting}
For CEX of compound OR: 1st sub-test must be a plain \asm{Bcc} (cond B in middle of IT illegal); only the FINAL test can fan out into a full \asm{ITTEE}.

\subsection{Cond LDR/STR forms in IT}
All three addressing modes (\asm{[Rn,\#os]}, \asm{[Rn,\#os]!}, \asm{[Rn],\#os}) accept a CCS suffix (Lctr 24 Ex 3,4): e.g.\ \asm{ldrshgt r0,[r2,\#2]!}, \asm{ldrle r0,[r3],\#4}. Pre/post writeback honored only when cond true. Cond instr that fails: no write, no flag change, no exception.

\subsection{HW11 P1 -- \cee{x\_cmp\_y\_load}, three styles}
\begin{hwp}\textbf{P1.} \cee{int32\_t f(int32\_t x, int32\_t y, int16\_t *ptrA, int32\_t *ptrB)\{ if (x>y) return (int32\_t)*(++ptrA)*2; else return *(ptrB++)/2; \}}\\
AAPCS: R0=x, R1=y, R2=ptrA, R3=ptrB; ret in R0.\\
Then arm: \cee{*(++ptrA)} $\to$ \asm{ldrsh r0,[r2,\#2]!} (pre-idx, sign-ext halfword); \cee{*2} $\to$ \asm{lsl r0,r0,\#1}.\\
Else arm: \cee{*(ptrB++)} $\to$ \asm{ldr r0,[r3],\#4} (post-idx); signed \cee{/2} $\to$ \asm{asr r0,r0,\#1}.

\textbf{Part 1 -- PL (BGT, same CCS as C):}
\begin{lstlisting}
x_cmp_y_load_pl:
    cmp     r0, r1
    bgt     pl_then         @ x>y -> then
pl_else:                    @ else lays out FIRST
    ldr     r0, [r3], #4    @ *ptrB; ptrB++
    asr     r0, r0, #1      @ /2 signed
    b       pl_end
pl_then:
    ldrsh   r0, [r2, #2]!   @ ++ptrA; *ptrA (s-ext)
    lsl     r0, r0, #1      @ *2
pl_end:
    bx      lr
\end{lstlisting}

\textbf{Part 2 -- NL (BLE, inverse CCS, reads like C):}
\begin{lstlisting}
x_cmp_y_load_nl:
    cmp     r0, r1
    ble     nl_else         @ x<=y -> else (skip then)
nl_then:                    @ then falls through
    ldrsh   r0, [r2, #2]!
    lsl     r0, r0, #1
    b       nl_end
nl_else:
    ldr     r0, [r3], #4
    asr     r0, r0, #1
nl_end:
    bx      lr
\end{lstlisting}

\textbf{Part 3 -- CEX (\asm{ITTEE le}, no branches):}
\begin{lstlisting}
x_cmp_y_load_cex:
    cmp     r0, r1
    ittee   le              @ T=le,T=le,E=gt,E=gt
    ldrle   r0, [r3], #4    @ T1: else load (post-idx)
    asrle   r0, r0, #1      @ T2: else /2
    ldrshgt r0, [r2, #2]!   @ E1: then load (pre-idx)
    lslgt   r0, r0, #1      @ E2: then *2
    bx      lr
\end{lstlisting}
CCS in \asm{ITTEE} = \asm{le} (NL cond = inverse of C's \cee{>}). Each arm is 2 instrs $\Rightarrow$ \asm{ITTEE} (4 slots) is the natural fit. Cond pre-idx (\asm{ldrshgt!}) and post-idx (\asm{ldrle ,\#4}) both legal.
\end{hwp}

\subsection{HW11 P2 -- \cee{mathf} piecewise (3-way cascade)}
\begin{hwp}\textbf{P2.} \cee{f(x) = -10 if x<-5; +10 if x>=5; else 2x}.\\
\textbf{Part 1 (C):}
\begin{lstlisting}[language={}]
if (x < -5)      y = -10;
else if (x >= 5) y =  10;
else             y =  2*x;
\end{lstlisting}

\textbf{Part 2 -- CMP chain (NL at each arm):}
\begin{lstlisting}
mathf_cmp:
    cmp     r0, #-5         @ asm: cmn r0,#5
    bge     elseif          @ NL: x>=-5 skips 1st arm
    ldr     r0, =-10        @ x<-5 -> -10
    b       end
elseif:
    cmp     r0, #5
    blt     else_           @ NL: x<5 skips 2nd arm
    ldr     r0, =10         @ x>=5 -> +10
    b       end
else_:
    lsl     r0, r0, #1      @ 2*x
end:
    bx      lr
\end{lstlisting}
\textbf{Note:} \asm{cmp r0,\#-5} not directly encodable; assembler emits \asm{cmn r0,\#5} (flags = R0+5). Signed CCSs work unchanged.

\textbf{Part 3 -- CEX (chained \asm{ITT lt} + \asm{ITE ge}):}
\begin{lstlisting}
mathf_cex:
    cmp     r0, #-5
    itt     lt              @ T=lt, T=lt
    ldrlt   r1, =-10        @ T1: stash -10 in R1
    blt     end             @ T2: cond B *LAST slot only*
    @ fall-through: x >= -5
    cmp     r0, #5          @ fresh CMP, new flags
    ite     ge              @ T=ge, E=lt
    ldrge   r1, =10         @ T : x>=5, R1=10
    lsllt   r1, r0, #1      @ E : R1=2*x (uses ORIG x in R0)
end:
    mov     r0, r1          @ result -> ret reg
    bx      lr
\end{lstlisting}
\textbf{Key tricks:}
\begin{itemize}
\item \textbf{Stash result in R1} so R0 (=x) survives the 2nd CMP. \asm{lsllt r1,r0,\#1} reads original x.
\item \asm{blt end} is in \textbf{slot 2} of \asm{ITT} -- legal only because it's the \textbf{last} slot (Lctr 24 rule).
\item Without the cond B, fall-through after 1st IT would corrupt R1 in 2nd block.
\end{itemize}
\end{hwp}

\subsection{HW11 P3 -- compound OR \cee{x<=20 || x>=25}}
\begin{hwp}\textbf{P3.} \cee{int32\_t g(int x, uint16\_t *ptrA, uint32\_t *ptrB)\{ if (x<=20 || x>=25) return (int32\_t)*++ptrA*4; else return *++ptrB/4; \}}\\
AAPCS: R0=x, R1=ptrA, R2=ptrB; ret in R0.\\
Then arm: \cee{*++ptrA} $\to$ \asm{ldrh r0,[r1,\#2]!} (zero-ext); \cee{*4} $\to$ \asm{lsl r0,r0,\#2}.\\
Else arm: \cee{*++ptrB} $\to$ \asm{ldr r0,[r2,\#4]!}; UNSIGNED \cee{/4} $\to$ \asm{lsr r0,r0,\#2} (NOT \asm{ASR} -- ptrB is \cee{uint32\_t*}).

\textbf{Part 1 -- CMP short-circuit OR (both tests $\to$ same \cee{then}):}
\begin{lstlisting}
load_based_on_x_cmp:
    cmp     r0, #20
    ble     then            @ 1st OR true -> then
    cmp     r0, #25
    bge     then            @ 2nd OR true -> then
    b       else_           @ both false -> else
then:
    ldrh    r0, [r1, #2]!   @ ++ptrA; *ptrA (zero-ext)
    lsl     r0, r0, #2      @ *4
    b       end
else_:
    ldr     r0, [r2, #4]!   @ ++ptrB; *ptrB
    lsr     r0, r0, #2      @ /4 unsigned
end:
    bx      lr
\end{lstlisting}

\textbf{Part 2 -- CEX (1st test plain \asm{BGT}, 2nd test \asm{ITTEE ge}):}
\begin{lstlisting}
load_based_on_x_cex:
    cmp     r0, #20
    bgt     check25         @ x>20 -> 2nd test (plain B; mid-IT cond B illegal)
    @ x<=20: compound TRUE -> then arm inline
    ldrh    r0, [r1, #2]!
    lsl     r0, r0, #2
    bx      lr
check25:
    cmp     r0, #25
    ittee   ge              @ T=ge,T=ge,E=lt,E=lt
    ldrhge  r0, [r1, #2]!   @ T: then load
    lslge   r0, r0, #2      @ T: *4
    ldrlt   r0, [r2, #4]!   @ E: else load
    lsrlt   r0, r0, #2      @ E: /4
    bx      lr
\end{lstlisting}
\textbf{Why split:} 1st sub-test of OR can't sit inside an IT block -- it would need a cond B in a non-last slot. Only the FINAL test packs into \asm{ITTEE}. Both arms run 2 instrs $\Rightarrow$ \asm{ITTEE} (T,T,E,E) is exact fit.

\textbf{Trace:}
\begin{tabular}{@{}c l l@{}}
\toprule
x & path & arm \\
\midrule
15 & \asm{bgt} N (15$\le$20) & inline then \\
20 & \asm{bgt} N            & inline then \\
21 & \asm{bgt} Y, ge N      & ITTEE E (else) \\
24 & \asm{bgt} Y, ge N      & ITTEE E (else) \\
25 & \asm{bgt} Y, ge Y      & ITTEE T (then) \\
30 & \asm{bgt} Y, ge Y      & ITTEE T (then) \\
\bottomrule
\end{tabular}
\end{hwp}

\subsection{Picker + pitfalls}
\textbf{Pick CEX} when both arms $\le$ 4 instrs total; \textbf{PL/NL} when arms larger or need general branch; \textbf{CMP short-circuit} for OR; \textbf{CMP+De Morgan} for AND (branch to else on $!a$/$!b$); \textbf{CBZ/CBNZ} for null/loop guards (forward $\le$126B, R0-R7, no flag effect).
\textbf{Gotchas:} (1) HI/LS on signed (or GT/LE on unsigned) miscompares near sign bit -- match CCS to type. (2) LSR on negative = huge positive; use ASR for signed div. (3) LDRH on \cee{int16\_t} loses sign -- use LDRSH. (4) Cond B in non-last IT slot = assembler err. (5) CBZ inside IT = forbidden. (6) In CEX cascade, stash result in R1 so R0 (=x) survives next CMP. (7) \asm{cmp Rn,\#-imm} rewritten to \asm{cmn Rn,\#imm} (signed CCS unchanged). (8) Pre/post-idx LDR both legal in IT, writeback honored only when cond true.

\subsection{One-line summary cards}
\begin{itemize}
\item \textbf{PL:} \asm{cmp; B<C-cond> then; <else>; b end; then: <then>; end:}
\item \textbf{NL:} \asm{cmp; B<inv-cond> else; <then>; b end; else: <else>; end:}
\item \textbf{CEX:} \asm{cmp; ITTEE <inv-cond>; <else x2 with cond>; <then x2 with inv-cond>}
\item \textbf{CEX cascade:} \asm{cmp; ITT lt; <stash-T>; blt end; cmp; ITE ge; <stash-T>; <stash-E>; end: mov r0,r1}
\item \textbf{OR (CMP):} \asm{cmp; ble then; cmp; bge then; b else}
\item \textbf{OR (CEX):} 1st = plain \asm{Bcc} to \asm{checkN}, 2nd packs into \asm{ITTEE}.
\end{itemize}

% ----- 09_hw12_loops.tex -----
\section{HW12: Loops in ASM}

\subsection{C loops $\to$ ASM shapes (Lctr 26 \S26.1--26.2)}
\begin{tabular}{@{}l l l l@{}}
\textbf{loop} & \textbf{test} & \textbf{min iters} & \textbf{C $\to$ ASM idiom} \\
\asm{while(c)} & top  & 0 & priming-read style: dup body update \\
\asm{do\{\}while(c)} & bottom & 1 & body, then \asm{cmp}+cond.\,branch back \\
\asm{for(i;c;u)} & top & 0 & \asm{init} / lp: \asm{cmp}+exit / body / \asm{u} / \asm{b lp} \\
\asm{while(1)\{...break;\}} & inline & 0+ & single in-loop \asm{LDRB}, \asm{break} = \asm{b \_end} \\
\end{tabular}

\textbf{Label conventions:} \asm{\_lp} at the top (re-entry target); \asm{\_end} after the exit (target of \asm{break}/condition-fail). \textbf{Top-tested while $\ge 0$ iters; do-while $\ge 1$ iter.} \asm{for} = same shape as \asm{while} with explicit init+update slots.

\subsection{Plain \asm{while} (priming-read style)}
\begin{lstlisting}
@ ch = *src++;                    @ priming LDRB BEFORE label
@ while (ch) { body; ch = *src++; }
        ldrb r2, [r1], #1         @ priming read
fn_lp:  cbz  r2, fn_end           @ exit when ch==0
            @ ... body ...
            ldrb r2, [r1], #1     @ in-loop update (DUP of priming)
        b    fn_lp
fn_end: bx   lr
\end{lstlisting}
Two textual \asm{LDRB}s (priming + update). \asm{CBZ} on the freshly-loaded byte tests the null-terminator without setting flags.

\subsection{\asm{do-while} (bottom-tested, body $\ge 1$)}
\begin{lstlisting}
@ do { body; i++; } while (i < N);
        mov  r1, #0               @ i = 0
fn_lp:                            @ no entry test -> body always runs once
            @ ... body ...
            add  r1, r1, #1
        cmp  r1, r2
        blt  fn_lp                @ back-edge test at bottom
fn_end: bx   lr
\end{lstlisting}
No \asm{\_end} jump-around needed; first entry falls straight into body.

\subsection{\asm{for} loop}
\begin{lstlisting}
@ for (i = 0; i < N; i++) { body }
        mov  r1, #0               @ INIT
fn_lp:  cmp  r1, r2               @ TEST at top
        bge  fn_end               @ exit if !cond
            @ ... body ...
            add  r1, r1, #1       @ UPDATE
        b    fn_lp
fn_end: bx   lr
\end{lstlisting}

\subsection{\asm{while(1)} + \asm{break} (cleaner than priming)}
\begin{lstlisting}
@ while (1) { ch = *src++; if (!ch) break; body; }
fn_lp:  ldrb r2, [r1], #1         @ SINGLE in-loop read (no priming)
        cbz  r2, fn_end           @ break = forward b end
            @ ... body ...
        b    fn_lp
fn_end: bx   lr
\end{lstlisting}
\textbf{Dynamic-count rule:} priming-read and \asm{while(1)+break} execute the \emph{same} number of \asm{LDRB}s at runtime ($N{+}1$ for an $N$-element string). The break-style only reduces \emph{source-line} count by removing the duplicated \asm{LDRB}.

\subsection{\asm{while}-with-CEX (rare; body must fit IT)}
\begin{lstlisting}
fn_lp:  cmp  r1, r2
        itt  lt
        ldrlt r0, [r3], #4        @ body in IT
        blt  fn_lp                @ cond B legal ONLY as last IT slot
\end{lstlisting}
Only viable when the entire body (and the back-edge \asm{B}) fits in 4 IT slots. Otherwise use plain \asm{cmp}+\asm{bcc}.

\subsection{\asm{break} / \asm{continue} semantics (Lctr 27 \S27.2--27.3)}
\textbf{break} $\to$ unconditional forward \asm{b \_end}; must be wrapped in an \asm{if} so it's executed conditionally (typically via \asm{CBZ}/\asm{CBNZ}/\asm{Bcc}).

\textbf{\asm{continue} target depends on loop type:}
\begin{tabular}{@{}l l@{}}
in \asm{while}/\asm{do-while} & jumps to \textbf{condition test} (\asm{\_lp}) \\
in \asm{for}                  & jumps to the \textbf{UPDATE expr} (i++ still runs) \\
\end{tabular}
$\Rightarrow$ in a while loop, any loop-control update must appear \emph{before} \asm{continue}; otherwise the loop hangs.

\subsection{HW12 P1 -- swap last 2 of every trio}
\begin{hwp}\textbf{P1.} \cee{int mp\_swap\_last\_2\_chars\_in\_3\_in\_str(char *str)} -- walk 3 chars at a time; if all 3 non-null, swap the last two; return \# trios swapped.\\
\textbf{Test:} \cee{"hellow"$\to$"hlelwo"} ret 2; \cee{"hello"$\to$"hlelo"} ret 1; \cee{"he"} ret 0.

\textbf{C (while(1)+break, three-way exit):}
\begin{lstlisting}
int n = 0;
while (1) {
    char c1 = *str++; if (!c1) break;
    char c2 = *str++; if (!c2) break;
    char c3 = *str++; if (!c3) break;
    *(str-2) = c3;  *(str-1) = c2;  n++;
}
return n;
\end{lstlisting}

\textbf{ASM} (\asm{R12} accumulator $\Rightarrow$ no push/pop; \asm{r1/r2/r3} = c1/c2/c3 scratch):
\begin{lstlisting}
        mov  r12, #0               @ pair_swapped = 0
fn_lp:  ldrb r1, [r0], #1          @ c1 = *str++  (post-idx)
        cbz  r1, fn_end            @ break if !c1
        ldrb r2, [r0], #1          @ c2 = *str++
        cbz  r2, fn_end            @ break if !c2
        ldrb r3, [r0], #1          @ c3 = *str++
        cbz  r3, fn_end            @ break if !c3
            strb r3, [r0, #-2]     @ *(str-2) = c3
            strb r2, [r0, #-1]     @ *(str-1) = c2
            add  r12, r12, #1      @ n++
        b    fn_lp
fn_end: mov  r0, r12               @ return n
        bx   lr
\end{lstlisting}

\textbf{Why \asm{[r0,\#-2]}/\asm{[r0,\#-1]} (not $-3,-2$):} after 3 post-inc \asm{LDRB}s, \asm{r0} sits one byte \emph{past} the trio. The trio occupies addrs \asm{[r0,\#-3]} (c1), \asm{[r0,\#-2]} (c2), \asm{[r0,\#-1]} (c3). We swap \emph{only the last two} $\Rightarrow$ $-2$ and $-1$. Using $-3,-2$ corrupts c1.

\textbf{Why no \asm{!} writeback on the STRBs:} we want \asm{r0} to remain at the post-trio cursor for the next iter; \asm{[r0,\#-2]} is plain base+neg-offset, no update.
\end{hwp}

\subsection{HW12 P2/P1 -- lower$\to$upper, plain \asm{while}}
\begin{hwp}\textbf{P2.1.} \cee{void mp\_str\_cpy\_2\_caps\_while(char *dst,char *src);} -- ASM of the given priming-read C.\\
\textbf{C (given):}
\begin{lstlisting}
char ch = *src++;          // priming read
while (ch) {
    ch -= 0x20;            // 'a'=0x61 -> 'A'=0x41
    *dst++ = ch;
    ch = *src++;           // duplicate update
}
*dst++ = ch;               // post-loop store of '\0'
\end{lstlisting}

\textbf{Regs (AAPCS):} r0=dst, r1=src, r2=ch. All scratch $\Rightarrow$ no push/pop.

\textbf{ASM:}
\begin{lstlisting}
        ldrb r2, [r1], #1          @ priming read
fn_lp:  cbz  r2, fn_end            @ while (ch)
            sub  r2, r2, #0x20     @ ch -= 0x20
            strb r2, [r0], #1      @ *dst++ = ch
            ldrb r2, [r1], #1      @ ch = *src++  (DUP)
        b    fn_lp
fn_end: strb r2, [r0], #1          @ *dst++ = '\0'  (r2=0 from CBZ exit)
        bx   lr
\end{lstlisting}
\textbf{Note:} \asm{0x20} is a Thumb-2 8-bit-rotated immediate $\Rightarrow$ \asm{sub} encodes directly. Terminator store at \asm{\_end} uses \asm{r2=0} guaranteed by \asm{CBZ}-exit. Writeback (\asm{\#1}) on terminator \asm{strb} is harmless -- \asm{r0} is dead.
\end{hwp}

\subsection{HW12 P2/P2 -- refactor to \asm{while(1)}+\asm{break} (C only)}
\begin{hwp}\textbf{P2.2.} Eliminate the duplicate \asm{*src++} read.
\begin{lstlisting}
while (1) {
    char ch = *src++;
    if (ch == 0) {
        *dst++ = ch;       // store '\0' BEFORE break
        break;
    }
    ch -= 0x20;
    *dst++ = ch;
}
\end{lstlisting}
Single read at top of body; \asm{break} after writing the terminator.\\
\textbf{Equivalence:} non-null bytes $\to$ read,sub,store; null byte $\to$ store-zero,exit. Identical observable behavior to plain \asm{while} version.
\end{hwp}

\subsection{HW12 P2/P3 -- \asm{while(1)}+\asm{break} ASM}
\begin{hwp}\textbf{P2.3.} ASM of the refactored C.
\begin{lstlisting}
fn_lp:  ldrb r2, [r1], #1          @ ch = *src++  (SINGLE read)
        cbz  r2, fn_end            @ if (!ch) goto end (deferred terminator)
            sub  r2, r2, #0x20
            strb r2, [r0], #1
        b    fn_lp
fn_end: strb r2, [r0], #1          @ *dst++ = '\0'  (r2=0 by CBZ)
        bx   lr
\end{lstlisting}
\textbf{Where the terminator store lives:} the C writes \asm{*dst++ = ch} \emph{inside} the \asm{if(!ch)} branch \emph{before} \asm{break}. Translating literally, that store must still happen $\Rightarrow$ place the \asm{strb} as the \emph{first} instruction at \asm{\_end}. Don't duplicate it inside the loop.\\
\textbf{Source-line vs dynamic count:} only one textual \asm{ldrb} (vs two in P2/P1). Runtime: same $N{+}1$ \asm{LDRB} executions per call. The refactor saves \emph{source bytes}, not cycles.
\end{hwp}

\subsection{Loop dynamic-count rule (memorize)}
For a string of $N$ non-null chars then \asm{\textbackslash 0}: priming-read \asm{while} and \asm{while(1)+break} both execute exactly $N{+}1$ \asm{LDRB}s at runtime (one per byte read, including the terminator). \textbf{Refactor only collapses source duplication; it does not save instructions executed.}

\subsection{Pointer-idiom $\to$ ASM (byte) cheat-card}
\begin{tabular}{@{}l l l@{}}
\cee{x = *p++}      & \asm{ldrb r0, [r1], \#1}    & post-idx, writeback \\
\cee{*p++ = x}      & \asm{strb r0, [r1], \#1}    & post-idx, writeback \\
\cee{x = *++p}      & \asm{ldrb r0, [r1, \#1]!}   & pre-idx, writeback \\
\cee{*++p = x}      & \asm{strb r0, [r1, \#1]!}   & pre-idx, writeback \\
\cee{*(p-1) = x}    & \asm{strb r0, [r1, \#-1]}   & base+neg-offset, no writeback \\
\cee{*(p-2) = x}    & \asm{strb r0, [r1, \#-2]}   & base+neg-offset, no writeback \\
\end{tabular}

\subsection{\asm{CBZ}/\asm{CBNZ} -- when, why, restrictions (Lctr 25)}
\textbf{\asm{CBZ Rn,lbl}} = $\{$\asm{cmp Rn,\#0}; \asm{beq lbl}$\}$ in one instr; \textbf{leaves flags untouched}.
\begin{itemize}
\item Ideal for null-terminator test on a fresh \asm{LDRB} (no flags wanted).
\item \asm{Rn} restricted to \asm{r0..r7}.
\item Forward branch only ($\le {+}126$ B); \emph{cannot} branch backward.
\item \textbf{Cannot appear inside an IT block} (Lctr 24 \S24.3.2).
\end{itemize}

\subsection{\asm{switch} implementations (Lctr 27 \S27.4)}
HW12 doesn't exercise \asm{switch}, but it's exam-fair. Three forms:

\textbf{(a) \asm{CMP}/\asm{BEQ} ladder} -- $O(n)$, one cmp per case:
\begin{lstlisting}
        cmp r0, #0
        beq case0
        cmp r0, #1
        beq case1
        ...
        b   default
\end{lstlisting}

\textbf{(b) Stacked-label fall-through} -- multiple values share a body by stacking labels:
\begin{lstlisting}
case0:
case1:                          @ both 0 and 1 fall here
        @ shared body
        b   end_sw
\end{lstlisting}

\textbf{(c) \asm{TBB} jump table} -- $O(1)$ via byte offsets:
\begin{lstlisting}
        adr  r1, jt
        tbb  [r1, r0]           @ PC += 2 * jt[r0]
jt:     .byte (case0 - jt)/2
        .byte (case1 - jt)/2
        .byte (case2 - jt)/2
        .byte (default - jt)/2
\end{lstlisting}
\textbf{\asm{TBH}} variant uses halfword offsets for far cases: \asm{tbh [r1, r0, lsl \#1]}, \asm{.short} entries. Use TBB for tight tables ($\le 510$ B reach), TBH for larger ones.

\subsection{Common HW12 mistakes}
\begin{itemize}
\item Wrong offsets in P1: \asm{\#-3,\#-2} instead of \asm{\#-2,\#-1} -- corrupts c1.
\item Putting the terminator \asm{strb} \emph{inside} the loop in P2/P3 -- duplicates the C's \cee{*dst++=ch;break} pattern; cleanest is at \asm{\_end}.
\item Using \asm{cmp Rn,\#0}+\asm{beq} when \asm{cbz} suffices -- wastes 2 bytes / blows away flags you may need.
\item Pushing \asm{r4} for a tiny loop accumulator -- prefer \asm{r12} (IP, scratch under AAPCS).
\item Trying \asm{cbz} inside an IT block -- forbidden.
\item Forgetting that \cee{while(ch)} on an empty input runs body 0 times -- the post-loop \cee{*dst++=ch} still writes the \asm{\textbackslash 0} (correct).
\item For-loop \asm{continue} branches to UPDATE, not the test -- so \asm{i++} still runs. While-loop \asm{continue} branches to test.
\end{itemize}

\subsection{Quick-reference summary}
\begin{itemize}
\item \textbf{Top-tested} (\asm{while}, \asm{for}) $\ge 0$ iters; \textbf{bottom-tested} (\asm{do-while}) $\ge 1$.
\item \textbf{Priming \asm{while}} = duplicated \asm{LDRB} (text); \textbf{\asm{while(1)+break}} = single \asm{LDRB}; same dynamic count.
\item Labels: \asm{\_lp} (top, re-entry) and \asm{\_end} (post-exit, terminator-store target).
\item \asm{CBZ} on fresh \asm{LDRB} = idiomatic null-terminator test; not allowed in IT.
\item Post-idx \asm{ldrb [r0],\#1} $=$ \cee{*p++}; base+neg-offset \asm{[r0,\#-k]} $=$ \cee{*(p-k)} no update.
\item \asm{R12} (IP) is AAPCS scratch -- prefer over \asm{R4} for short-lived loop locals.
\item \asm{break} $\to$ forward \asm{b \_end}; \asm{continue} target = \asm{for}'s update / \asm{while}'s test.
\item \asm{switch}: ladder ($O(n)$) / stacked fall-through (shared body) / \asm{TBB},\asm{TBH} jump table ($O(1)$).
\end{itemize}

% ----- 10_quiz3.tex -----
% =============================================================================
% Quiz 3 --- early-ASM material (EABI, stack, barrel-shifter, BIC+ORR splice)
% NOTE: no standalone qz3 problem/solution PDF exists in the course repo. The
% problems below are reconstructed from
%   content/cec_320/quizes/quiz3/qz3_cheatsheet.md
%   content/cec_320/quizes/quiz3/qz3_init.md
%   content/cec_320/quizes/quiz3/lecture_notes_summary.md (Lctr 14--18)
%   content/cec_320/final_exam/quiz3_quiz4_extracts.md (Q3 P1..P6)
% Header in qz3_init.md is explicit: ``Just a reminder that you can use your own
% formula sheets for QZ3'' --- so the user originally also requested NVIC/EXTI
% material here, but Quiz 3 actually covers HW6/7/8 (Lctr 14--19): mixed C/ASM,
% EABI, arithmetic, shifts, bitwise logic, NZCV, CCS. No interrupt/EXTI/NVIC
% content was tested on Quiz 3 per the source markdown.
% =============================================================================
\section{Quiz 3: Interrupts + Early ASM}

\subsection{Coverage \& meta}
\textbf{Scope:} HW6 (mixed C+ASM, EABI, stack, 64-bit \asm{ADDS/ADC}, \asm{SUBS/SBC}, \asm{UDIV+MLS} modulo, Q15.16 encode), HW7 (shifts, barrel-shifter scaling, BIC/ORR bit-field splice), HW8 (NZCV, CCS, neg-logic branch, \asm{SMLAL} loop). Lectures 14--19. \emph{Despite the section title above, no NVIC/EXTI was on Q3 ---} that material lives on Q4/Q5. The ``early ASM'' tag is the accurate description; cross-reference \S HW6/7/8 above for full mechanics.\\
\textbf{Allowed:} prior cheat-sheets + one new letter-size sheet, only ASM instructions covered in lecture, no invented mnemonics.\\
\textbf{Reconstruction note:} Q3 problem statements below are derived from the cheatsheet+extract markdown; the original Fall-2025 PDF is not in the repo, so wording may differ slightly. Numeric answers are verified against the cheatsheet trace.

\subsection{One-page Q3 cookbook}
\begin{tabular}{@{}ll@{}}
\toprule
task & one-liner \\
\midrule
return signed $-2$ in 32b   & \cee{R0 = 0xFFFF\_FFFE} (two's-comp), \emph{not} 0x80000002 \\
sign-ext \cee{int8\_t} arg  & MSB$=$1 $\Rightarrow$ pad upper 24b with 1s; MSB$=$0 $\Rightarrow$ pad with 0s \\
zero-ext \cee{uint8\_t}     & always pad upper bits with 0s \\
\cee{push \{list\}} layout  & lowest reg\# at lowest addr (independent of brace order) \\
\cee{push} of $n$ regs      & \cee{SP -= 4*n} (full-descending) \\
$k\cdot B/2^n$ in 1 instr   & decompose $k/2^n=1\pm 1/2^m$, use \cee{ADD/SUB Rd,B,B,ASR\#m} \\
clear field then insert     & \cee{BIC Rd,Rd,mask\,LSL\#s}; \cee{ORR Rd,Rd,val\,LSL\#s} \\
\cee{ADDS/ADC} 64-bit add   & low \cee{ADDS} sets C; \cee{ADC} adds C into high \\
\bottomrule
\end{tabular}

\subsection{P1 (7pt) --- Return reg for signed int32}
\textbf{Setup:} ASM function returns signed 32-bit $-2$. (a) Which register holds it? (b) Hex value?
\begin{qp}
(a) \textbf{R0} (AAPCS: any 32-bit return goes in R0; 64-bit return uses R0:R1 lo:hi). (b) $-2$ in 32-bit two's-complement: invert bits of $+2 = $\cee{0x0000\_0002} and add 1 $\to$ \textbf{\cee{0xFFFF\_FFFE}}. \emph{Trap:} writing \cee{0x8000\_0002} (sign-magnitude form, which ARM does \emph{not} use) loses 6 of 7 points.
\end{qp}

\subsection{P2 (20pt) --- 64-bit ADD + C flag}
\textbf{Setup:} \asm{mp\_add\_64bit\_s} executes \asm{adds r0,r2}; \asm{adc r1,r3}; \asm{bx lr}. Two test sets: $A_1=$\cee{0x377{<}{<}24}, $B_1=$\cee{0x287{<}{<}24}; $A_2=$\cee{0x378{<}{<}24}, $B_2=$\cee{0x288{<}{<}24}. Find sums and the C flag observed after the low-word \asm{adds}.
\begin{qp}
Layout: \asm{R1:R0} holds A, \asm{R3:R2} holds B. Bits 24..31 of the low word receive the most-significant nibble of the constant.\\
\textbf{Set 1:} low words \cee{0x77\_0000\_00 + 0x87\_0000\_00 = 0xFE\_0000\_00} (no carry-out of bit 31) $\Rightarrow$ \textbf{C=0}. High words 0+0+0=0. Sum $=$ \cee{(int64\_t)0x5FE{<}{<}24}.\\
\textbf{Set 2:} low \cee{0x78\_0000\_00 + 0x88\_0000\_00 = 0x1\_00\_0000\_00} (overflows bit 31!) $\Rightarrow$ \textbf{C=1}, low result \cee{0x00\_0000\_00}. \asm{adc} adds C to high: $0+0+1=1$. Sum $=$ \cee{(int64\_t)0x600{<}{<}24}.\\
\emph{Design rule:} a good 64-bit-ADD test toggles the low-word carry between cases (one set with C=0, one with C=1) so \asm{adc} is exercised both ways.
\end{qp}

\subsection{P3 (20pt) --- \texttt{push \{r4,r8,r6\}} memory layout}
\textbf{Setup:} $R_i = (1{<}{<}i) + (1{<}{<}(2{+}i))$ for $i=1..12$. \asm{SP} $=$ \cee{0x2000\_1020} before \asm{push \{r4, r8, r6\}}. Find SP after, $R_4$, the word at \cee{0x2000\_1018}, and the byte at \cee{0x2000\_1019}.
\begin{qp}
3 regs $\cdot$ 4B $=$ 12 $=$ \cee{0xC} $\Rightarrow$ \textbf{SP$=$\cee{0x2000\_1014}}. Push writes \emph{in ascending register-number order} regardless of brace listing $\Rightarrow$ R4 lowest, R6 middle, R8 highest:\\
\quad \cee{0x2000\_1014} $\leftarrow$ \textbf{R4} = \cee{(1{<}{<}4)+(1{<}{<}6) = 0x10+0x40 = 0x0000\_0050}\\
\quad \cee{0x2000\_1018} $\leftarrow$ \textbf{R6} = \cee{(1{<}{<}6)+(1{<}{<}8) = 0x40+0x100 = 0x0000\_0140}\\
\quad \cee{0x2000\_101C} $\leftarrow$ \textbf{R8} = \cee{(1{<}{<}8)+(1{<}{<}10) = 0x100+0x400 = 0x0000\_0500}\\
Word at \cee{0x2000\_1018} $=$ \textbf{\cee{0x0000\_0140}}. Byte at \cee{0x2000\_1019}: little-endian, address +1 holds byte 1 of R6 $=$ \textbf{\cee{0x01}} (next bytes: \cee{0x00},\cee{0x00},\cee{0x00} for the upper word).\\
\emph{Trap:} writing in brace-list order (R4,R8,R6) and putting R8 at the middle address --- the assembler ignores the listing order.
\end{qp}

\subsection{P4 (25pt) --- Single-instr $k\cdot B/8$ via barrel shifter}
\textbf{Setup:} signed $A$ in \asm{r0}, signed $B$ in \asm{r1}. Each part wants $A=k\cdot B/8$ in one instruction. Decompose as $k/8 = 1 \pm 2^{-m}$ where convenient; remainder $= 1$ becomes a no-op MOV.
\begin{center}\scriptsize
\begin{tabular}{@{}clll@{}}
\toprule
$k$ & $k/8$ & decomp & instruction \\
\midrule
6  & 6/8  & $1 - 2^{-2}$ & \asm{sub r0, r1, r1, asr \#2} \\
7  & 7/8  & $1 - 2^{-3}$ & \asm{sub r0, r1, r1, asr \#3} \\
8  & 1    & $1$          & \asm{mov r0, r1} \;\,(or \asm{add r0, r1, \#0}) \\
9  & 9/8  & $1 + 2^{-3}$ & \asm{add r0, r1, r1, asr \#3} \\
10 & 10/8 & $1 + 2^{-2}$ & \asm{add r0, r1, r1, asr \#2} \\
\bottomrule
\end{tabular}
\end{center}
\begin{qp}
\textbf{Why ASR not LSR?} $B$ is signed; \asm{LSR} fills MSB with 0, mangling negatives ($-1{\,>>\,}1$ would become $\!\sim\!2^{31}$). \asm{ASR} replicates the sign bit. \textbf{Why no \asm{rsb}?} $1\pm 2^{-m}$ keeps the unshifted $B$ first, so the sign of $B$ leads the result; \asm{rsb} would only be needed for $-1+\text{something}$ patterns ($k\cdot B$ with $k$ near 0). \textbf{Trap:} \asm{add r0, r0, r0, asr \#3} silently corrupts when \asm{r0=r1} aliases the input.
\end{qp}

\subsection{P5 (20pt) --- BIC+ORR bit-field splice $\to$ C}
\textbf{Setup:} ASM body
\begin{lstlisting}
mp_func_s:
    ldr  r3, =15           @ mask = 0xF (4 ones)
    bic  r0, r0, r3, lsl #8   @ a &= ~(0xF << 8)   ; clear bits[11:8]
    and  r1, r1, r3           @ b &= 0xF           ; isolate b's low nibble
    orr  r0, r0, r1, lsl #8   @ a |= b << 8        ; insert into [11:8]
    bx   lr
\end{lstlisting}
Write the C equivalent and trace with $a=$\cee{0xH3H2H1H0}, $b=17=$\cee{0x0011}.
\begin{qp}
C version (canonical clear-then-insert):
\begin{lstlisting}[language=]
uint16_t func_c(uint16_t a, uint16_t b) {
    uint16_t mask = 15;
    a &= ~(mask << 8);          // clear nibble at bit 8
    a |= (b & mask) << 8;       // insert low nibble of b
    return a;
}
\end{lstlisting}
\textbf{Trace} with $a=$\cee{0xH3H2H1H0}, $b=$\cee{0x0011}: \asm{r3=0x000F}; \asm{r3{<}{<}8 = 0x0F00}; \asm{bic} $\Rightarrow a=$\cee{0xH3\_0\_H1H0} (digit H2 cleared). $b\,\&\,0xF = 0x1$; \asm{orr} with \cee{0x1{<}{<}8 = 0x0100} $\Rightarrow$ \textbf{$a=$\cee{0xH3\,1\,H1H0}}. \emph{Pattern alias:} this is the universal ``\asm{BIC}+\asm{ORR} both shifted by the same \asm{LSL\#s}'' field-splice idiom; if you see \asm{BIC ...,LSL\#s} look immediately for the matching \asm{ORR ...,LSL\#s} below.
\end{qp}

\subsection{P6 (8pt) --- \cee{int8\_t} sign-ext into 32-bit args}
\textbf{Setup:} \cee{int8\_t var1=0x7E, var2=0x81; fn(var1,var2);} where prototype takes \cee{int32\_t}. Find R0, R1.
\begin{qp}
Sign-extension occurs at the \emph{call site} per AAPCS, before the callee runs.\\
\cee{var1=0x7E}: MSB (bit 7) $=$ 0 $\Rightarrow$ positive ($+126$); pad upper 24b with 0 $\Rightarrow$ \textbf{R0$=$\cee{0x0000\_007E}}.\\
\cee{var2=0x81}: MSB $=$ 1 $\Rightarrow$ negative ($-127$ as \cee{int8\_t}); pad upper 24b with 1 $\Rightarrow$ \textbf{R1$=$\cee{0xFFFF\_FF81}}.\\
\emph{Sanity:} \cee{0xFFFF\_FF81} as \cee{int32\_t} is $-127$ --- value preserved across the widening. Compare to \cee{uint8\_t} promotion (always zero-ext): \cee{0x81} would become \cee{0x0000\_0081} ($=129$). The same sign-ext rule governs \asm{LDRSB} / \asm{LDRSH} loads (Q4/Q6a territory).
\end{qp}

\subsection{Cross-references --- prior-week material on Q3}
\textbf{EABI registers:} R0--R3 args/scratch (caller-saved); R4--R11 callee-saved; R12=IP (scratch); R13=SP, R14=LR, R15=PC. 64-bit return: R0=lo, R1=hi. \textbf{Stack:} full-descending, SP points to last pushed item, \asm{push} pre-decrements, \asm{pop} post-increments. \textbf{Flag-only ops:} \asm{CMP=SUBS}-discard, \asm{CMN=ADDS}-discard, \asm{TST=ANDS}-discard (no V update), \asm{TEQ=EORS}-discard. \textbf{C-on-sub:} ARM sets C$=$!borrow (opposite of x86). \textbf{CCS domains:} \asm{HI/HS/LO/LS} unsigned (use C+Z), \asm{GT/GE/LT/LE} signed (use V vs N), \asm{EQ/NE} domain-free (just Z). \textbf{Neg-logic branch:} after \asm{cmp}, branch on the \emph{opposite} CCS to skip the then-block.

\subsection{Q3 trap inventory (memorize before walking in)}
\begin{itemize}
  \item Negative int return $\to$ two's-complement hex, \emph{not} sign-magnitude (P1).
  \item \asm{push \{r4,r8,r6\}} stores in numeric order; brace-list order is irrelevant (P3).
  \item 64-bit add: low \asm{adds} provides C, high \asm{adc} consumes it; if your test only exercises one C value, \asm{adc} is half-untested (P2).
  \item Barrel-shifter ASR vs LSR --- ASR for signed, LSR for unsigned. Mismatch $=$ silent corruption (P4).
  \item \asm{BIC \dots LSL\#s} immediately followed by \asm{ORR \dots LSL\#s} $=$ ``replace field [s+w-1:s]'' --- treat as a single semantic unit (P5).
  \item \cee{int8\_t} args $\geq$ \cee{0x80} sign-extend to \cee{0xFFFF\_FFxx}; \cee{uint8\_t} always zero-extends. Same rule for \asm{ldrsb} on Q4 (P6).
  \item C $=$ NOT borrow on subtract (ARM convention). On 5-bit hand-flag drills, this flips most students up.
  \item \asm{TST}/\asm{TEQ} don't update V $\to$ never pair with V-dependent CCS.
\end{itemize}

% ----- 11_quiz4_qz26a.tex -----
% =============================================================================
% Quiz 4 (qz-26a, Spring 2026) + practice qz-25b qz5/qz6a
% Topics: LDR/STR (signed/half/byte), pre/post-index +writeback, scaled-reg
%         addressing, LDRD/STRD, conditional-branch flow control, ptr idioms.
% =============================================================================
\section{Quiz 4: LDR/STR + Flow Control (qz-26a current sem)}

\subsection{Cheat-card: LDR/STR variants}
\textbf{Sign-extend loads} fill upper bits with the loaded MSB:
\begin{center}\scriptsize
\begin{tabular}{@{}llll@{}}
\toprule
mnem & loads & upper bits & view \\
\midrule
\asm{LDRB}  & 1 byte  & zero-fill   & unsigned \\
\asm{LDRSB} & 1 byte  & sign-extend & signed   \\
\asm{LDRH}  & 2 bytes & zero-fill   & unsigned \\
\asm{LDRSH} & 2 bytes & sign-extend & signed   \\
\asm{LDR}   & 4 bytes & ---         & word     \\
\asm{LDRD}  & 8 bytes & Rt=lo,Rt2=hi& pair     \\
\bottomrule
\end{tabular}
\end{center}
Stores have no signed variant (\asm{STRB/STRH/STR/STRD} only) -- low bits go to memory regardless of sign.

\textbf{Addressing modes} (effective addr {[}EA{]} vs writeback to Rn):
\begin{center}\scriptsize
\begin{tabular}{@{}lll@{}}
\toprule
syntax & EA & Rn after \\
\midrule
\asm{[Rn]}                & Rn          & Rn (unchanged) \\
\asm{[Rn,\#i]}            & Rn$+$i      & Rn (offset) \\
\asm{[Rn,\#i]!}           & Rn$+$i      & Rn$+$i \,\,(\textbf{pre-index} wb) \\
\asm{[Rn],\#i}            & Rn          & Rn$+$i \,\,(\textbf{post-index} wb) \\
\asm{[Rn,Rm,LSL \#n]}     & Rn$+$(Rm$\!\ll\!$n) & Rn (scaled-reg) \\
\bottomrule
\end{tabular}
\end{center}
\textbf{Pre} uses updated addr; \textbf{post} uses old addr then advances. Post-index by 4 advances Rn even if only 2 bytes were read (e.g.\ \asm{ldrsh r3,[r0],\#4}).

\textbf{Scaled register}: \asm{LDR Rt,[Rn,Rm,LSL \#n]} where \asm{n}$=\log_2$sizeof. \asm{LSL \#0} byte, \asm{\#1} half, \asm{\#2} word, \asm{\#3} dword. R1 is in \emph{elements}; the shift converts to bytes. Halfword/word access at non-aligned addr faults.

% =============================================================================
\subsection{Q4 (qz-26a, current sem)}
\textbf{Memory state for Prob 1} (little-endian; bytes labeled B0..B3):
\begin{center}\scriptsize
\begin{tabular}{@{}lcccc@{}}
\toprule
Addr & B0 & B1 & B2 & B3 \\
\midrule
0x2000\_0000 & 88 & 89 & 8A & 8B \\
0x2000\_0004 & 8C & 8D & 8E & 8F \\
0x2000\_0008 & 70 & 71 & 72 & 73 \\
0x2000\_000C & 74 & 75 & 76 & 77 \\
\bottomrule
\end{tabular}
\end{center}
Half/word reads are LE: addr 0x4 halfword $=$ \asm{0x8D8C} (B1:B0); word $=$ \asm{0x8F8E\_8D8C}.

\begin{qp}\textbf{Prob 1a (8pt) -- task1 just before \asm{bx lr}:}
\begin{lstlisting}
task1:                       @ R0=0x2000_0000, R1=0x2000_0020
    ldrsh r2, [r0, #4]       @ HW @0x4 = 0x8D8C (neg)
                             @ -> R2 = 0xFFFF_8D8C
    str   r2, [r1], #4       @ store full r2 to 0x20, R1+=4
    ldrsh r3, [r0, #8]       @ HW @0x8 = 0x7170 (pos)
                             @ -> R3 = 0x0000_7170
    str   r3, [r1], #4       @ store to 0x24, R1+=4
    bx    lr
\end{lstlisting}
\textbf{Ans:} \asm{R2=0xFFFF\_8D8C}, \asm{R3=0x0000\_7170}.\\
\emph{Trap:} \asm{LDRSH} with positive HW (MSB=0) gives upper-16 zeros; negative HW (MSB=1) gives upper-16 ones. R0 is \textbf{not} modified (no \asm{!}).
\end{qp}

\begin{qp}\textbf{Prob 1b (8pt) -- task2 (pre-index w/ writeback):}
\begin{lstlisting}
task2:                       @ R0=0x2000_0000, R1=0x2000_0030
    ldrsb r2, [r0, #2]!      @ R0 first becomes 0x2000_0002
                             @ load byte 0x8A, sign-ext
                             @ -> R2 = 0xFFFF_FF8A
    str   r2, [r1, #0]       @ store full r2 to 0x30
    ldrsb r3, [r0, #4]!      @ R0 -> 0x2000_0006
                             @ byte 0x8E, sign-ext
                             @ -> R3 = 0xFFFF_FF8E
    str   r3, [r1, #4]       @ store full r3 to 0x34
    bx    lr
\end{lstlisting}
\textbf{Ans:} \asm{R2=0xFFFF\_FF8A}, \asm{R3=0xFFFF\_FF8E}.
After return: \asm{R0=0x2000\_0006} (writeback persists across both pre-index ops).
\end{qp}

\begin{qp}\textbf{Prob 1c (8pt) -- mem map at 0x2000\_0020 after task1:}
task1 stores full \emph{word} R2 then word R3 with post-index. R2$=$\asm{0xFFFF\_8D8C} $\to$ bytes \asm{8C 8D FF FF}; R3$=$\asm{0x0000\_7170} $\to$ \asm{70 71 00 00}.
\begin{center}\scriptsize
\begin{tabular}{@{}lcccc@{}}
\toprule
Addr & B0 & B1 & B2 & B3 \\
\midrule
0x2000\_0020 & 8C & 8D & FF & FF \\
0x2000\_0024 & 70 & 71 & 00 & 00 \\
\bottomrule
\end{tabular}
\end{center}
\emph{LE rule:} R[7:0]$\to$B0, R[15:8]$\to$B1, R[23:16]$\to$B2, R[31:24]$\to$B3.
\end{qp}

\begin{qp}\textbf{Prob 1d (6pt) -- C \cee{val=*(ptr+i)} for \cee{uint16\_t *ptr}:}
\begin{lstlisting}
LDRH R2, [R0, R1, LSL #1]   @ scale i by 2 (sizeof uint16_t)
\end{lstlisting}
\emph{Trap:} forgetting \asm{LSL \#1} treats i as a byte offset $\Rightarrow$ wrong elem and possible alignment fault. For \cee{uint8\_t*} use \asm{LSL \#0} (or omit shift); \cee{int32\_t*} use \asm{LSL \#2}.
\end{qp}

\begin{qp}\textbf{Prob 2 (30pt) -- C ptr ops $\to$ ASM + memory.}\\
Initial mem (LE):
\begin{center}\scriptsize
\begin{tabular}{@{}lcccc@{}}
\toprule
Addr & B0 & B1 & B2 & B3 \\
\midrule
0x2000\_0000 & 7C & 7D & 7E & 7F \\
0x2000\_0004 & 80 & 81 & 82 & 83 \\
0x2000\_0008 & 84 & 85 & 86 & 87 \\
0x2000\_000C & 88 & 89 & 8A & 8B \\
\bottomrule
\end{tabular}
\end{center}
\cee{int16\_t *src=0x2000\_0000} in R0; \cee{int16\_t *dst=0x2000\_0008} in R1.
\begin{lstlisting}
@ (1) *dst++ = *src++;       @ post-inc both
LDRSH R2, [R0], #2           @ R2 = 0x7D7C; R0 -> 0x...02
STRH  R2, [R1], #2           @ HW@0x08 = 7C 7D; R1 -> 0x...0A

@ (2) *dst++ = *++src;       @ pre-inc src, then deref
LDRSH R2, [R0, #2]!          @ R0 -> 0x...04; R2 = 0xFFFF_8180
STRH  R2, [R1], #2           @ HW@0x0A = 80 81; R1 -> 0x...0C

@ (3) *dst++ = ++*src;       @ inc *src in place, then store
LDRSH R2, [R0]               @ R2 = 0xFFFF_8180
ADD   R2, #1                 @ R2 = 0xFFFF_8181
STRH  R2, [R0]               @ writeback HW@0x04 = 81 81
STRH  R2, [R1], #2           @ HW@0x0C = 81 81; R1 -> 0x...0E
\end{lstlisting}
\textbf{Final memory:}
\begin{center}\scriptsize
\begin{tabular}{@{}lcccc@{}}
\toprule
Addr & B0 & B1 & B2 & B3 \\
\midrule
0x2000\_0000 & 7C & 7D & 7E & 7F \\
0x2000\_0004 & \textbf{81} & \textbf{81} & 82 & 83 \\
0x2000\_0008 & \textbf{7C} & \textbf{7D} & \textbf{80} & \textbf{81} \\
0x2000\_000C & \textbf{81} & \textbf{81} & 8A & 8B \\
\bottomrule
\end{tabular}
\end{center}
\emph{Idiom map:}
\begin{center}\scriptsize
\begin{tabular}{@{}ll@{}}
\toprule
C & ASM pattern \\
\midrule
\cee{*p++}    & post-index \asm{[Rn],\#sz} \\
\cee{*++p}    & pre-index  \asm{[Rn,\#sz]!} \\
\cee{++*p}    & load,\,ADD\,\#1,\,STORE-back to \asm{[Rn]} \\
\cee{(*p)++}  & load to Rt,\,ADD\,\#1 to copy Rs,\,STORE Rs \\
\bottomrule
\end{tabular}
\end{center}
\emph{Trap:} \cee{++*p} modifies the value at \cee{*p}, NOT the pointer. The pointer p is unchanged; only the storage cell is incremented. Don't forget the writeback \asm{STRH R2,[R0]}.
\end{qp}

\begin{qp}\textbf{Prob 3 (20pt) -- LDRD/STRD round-trip.}\\
Setup: \cee{ipt[16]} filled with \cee{ipt[i]=i<<4} (so 00,10,20,...,F0). \cee{ipt=0x2000\_0200}, \cee{opt=0x2000\_0280}.
\begin{lstlisting}
mp_task:                  @ R0=ipt, R1=opt
    ldrd r2,r3,[r0,#8]!   @ R0+=8 -> 0x208
                          @ R2 = word @0x208 = 0xB0A0_9080
                          @ R3 = word @0x20C = 0xF0E0_D0C0
    str  r3,[r1,#0]       @ opt[0] = 0xF0E0_D0C0
    str  r2,[r1,#4]       @ opt[1] = 0xB0A0_9080
    sub  r0,r0,#8         @ R0 -> 0x200 (rewind)
    ldrsh r3,[r0],#4      @ HW@0x200 = 0x1000 (pos)
                          @ R3 = 0x0000_1000; R0 -> 0x204
    ldrsh r2,[r0,#4]!     @ R0 -> 0x208; HW@0x208 = 0x9080
                          @ R2 = 0xFFFF_9080 (sign-ext)
    strd r2,r3,[r1,#8]!   @ R1+=8 -> 0x288
                          @ word@0x288 = 0xFFFF_9080 (Rt=lo)
                          @ word@0x28C = 0x0000_1000 (Rt2=hi)
    bx   lr
\end{lstlisting}
\textbf{Part (a, 12pt) -- printout:}
\begin{lstlisting}
Out1=0xF0E0D0C0, Out2=0xB0A09080,
Out3=0xFFFF9080, Out4=0x1000
\end{lstlisting}
\textbf{Part (b, 8pt) -- final regs:} \asm{R0=0x2000\_0208}, \asm{R1=0x2000\_0288}.\\
\emph{Trap:} \asm{LDRD Rt,Rt2,[Rn,\#i]!} loads Rt from EA (low addr) and Rt2 from EA$+$4. Pair must be \emph{consecutive} regs (Rt2$=$Rt$+$1) and Rt usually even (legacy A32 constraint -- T32 relaxed but exam expects pairs). \asm{LDRSH...,\#4} post-indexes Rn by 4 even though only 2 bytes were read.
\end{qp}

\begin{qp}\textbf{Prob 4 (25pt) -- conditional branch translation.}\\
C source:
\begin{lstlisting}
int load_based_on_x_c(int x, int *ptrA, int *ptrB){
  if (x <= -20 || x >= 20) return *(ptrA+1) * 4;
  else                     return *(ptrB+2) / 4;
}
\end{lstlisting}
ASM (cond-branch method; x$\to$R0, ptrA$\to$R1, ptrB$\to$R2):
\begin{lstlisting}
load_based_on_x_s:
    cmp  r0, #-20
    ble  load_based_on_x_s_then  @ x<=-20 short-circuit to then
    cmp  r0, #20
    blt  load_based_on_x_s_else  @ -20<x<20 -> else
load_based_on_x_s_then:
    ldr  r0, [r1, #4]            @ *(ptrA+1): int*, off=1*4=4
    lsl  r0, #2                  @ *4
    b    load_based_on_x_s_end
load_based_on_x_s_else:
    ldr  r0, [r2, #8]            @ *(ptrB+2): int*, off=2*4=8
    lsr  r0, #2                  @ /4 (signed: should be ASR)
load_based_on_x_s_end:
    bx   lr
\end{lstlisting}
\emph{Traps:}
(1) \cee{||} short-circuits: 1st condition true $\Rightarrow$ jump to then \emph{without} testing the 2nd; 1st false $\Rightarrow$ test 2nd; both false $\Rightarrow$ else.
(2) Pointer offset is \textbf{bytes}: \cee{*(p+1)} for \cee{int*} is \asm{\#4}; \cee{*(p+2)} is \asm{\#8}.
(3) Signed div by $2^n$ is \asm{ASR \#n}; the soln uses \asm{LSR} which is wrong for signed but matches the rubric. Memorize: \asm{ASR}=signed, \asm{LSR}=unsigned.
(4) The closing \asm{b ...\_end} is mandatory to skip the else-block when then ran.
\end{qp}

% =============================================================================
\subsection{Practice qz-25b qz5}
Mem for prob 3:
\begin{center}\scriptsize
\begin{tabular}{@{}lcccc@{}}
\toprule
Addr & B0 & B1 & B2 & B3 \\
\midrule
0x2000\_0000 & 80 & 81 & 82 & 83 \\
0x2000\_0004 & 84 & 85 & 86 & 87 \\
0x2000\_0008 & 78 & 79 & 7A & 7B \\
0x2000\_000C & 7C & 7D & 7E & 7F \\
\bottomrule
\end{tabular}
\end{center}

\begin{qp}\textbf{P1 (20pt) -- mp\_selector(a,x,y): return a$\geq$0 ? x : y.}
\begin{lstlisting}
@ a->r0, x->r1, y->r2 ; return -> r0
mp_selector:
    cmp  r0, #0
    blt  mp_selector_else  @ neg-logic: a<0 -> else
mp_selector_then:
    mov  r0, r1
    bx   lr
mp_selector_else:
    mov  r0, r2
    bx   lr
\end{lstlisting}
\emph{Trap:} for signed compare to 0 use \asm{BLT}; for unsigned use \asm{BLO}. Two \asm{bx lr} (one per branch) is OK -- saves the unconditional \asm{b end}.
\end{qp}

\begin{qp}\textbf{P2 -- 5-bit flag drills (ARM convention: C$=$\textbf{!borrow} on sub).}
Constants: $C_N{=}32$, $U_{\!max}{=}31$, $S_{min}{=}-16$, $S_{max}{=}15$.

\textbf{(1) $-15 - (-16)$:}
\begin{itemize}
\item Unsigned: $(17-16)=1\in[0,31]$ $\Rightarrow$ no borrow $\Rightarrow$ \textbf{C$=$1}, dec$=$1.
\item Signed: $1\in[-16,15]$ $\Rightarrow$ \textbf{V$=$0}, dec$=$1.
\item Bin $=$ \cee{0\,0001} $\Rightarrow$ \textbf{N$=$0, Z$=$0}. Final \textbf{NZCV $=$ 0010}.
\end{itemize}

\textbf{(2) $16 - 18$:}
\begin{itemize}
\item Unsigned: $16-18 = -2\notin[0,31]$ $\Rightarrow$ borrow $\Rightarrow$ \textbf{C$=$0}; correct $-2{+}32=30$.
\item Signed: re-encode operands as 5-bit signed: $16\to-16$, $18\to-14$; $(-16)-(-14)=-2\in[-16,15]$ $\Rightarrow$ \textbf{V$=$0}, dec$=-2$.
\item Bin $=$ \cee{1\,1110} $\Rightarrow$ \textbf{N$=$1, Z$=$0}. Final \textbf{NZCV $=$ 1000}.
\end{itemize}
\emph{Trap:} unsigned-overflow $\to$ ``correct by $\pm C_N$''; signed needs operands re-interpreted (operand $\geq C_N/2$ becomes operand$-C_N$). $16$ in 5-bit signed is unrepresentable as $+16$, so it pattern-aliases to $-16$.
\end{qp}

\begin{qp}\textbf{P3 -- LDR variants \& scaled register.}
\begin{lstlisting}
task1:                       @ R0=0x2000_0000, R1=0x2000_0020
    ldr  r2, [r0, #4]!       @ R0->0x04; R2=word@0x04=0x8786_8584
    str  r2, [r1, #0]
    ldr  r3, [r0, #4]!       @ R0->0x08; R3=word@0x08=0x7B7A_7978
    str  r3, [r1, #4]
    bx   lr

task2:                       @ R0=0x2000_0000, R1=0x2000_0030, R2=-1
    ldr  r3, [r0], #8        @ R3=word@0x00=0x8382_8180; R0->0x08
    str  r3, [r1, #0]
    ldr  r3, [r0, r2, lsl #2]@ EA = 0x08 + (-1<<2) = 0x04
                             @ R3 = word@0x04 = 0x8786_8584
    str  r3, [r1, #4]
    bx   lr
\end{lstlisting}
\textbf{Ans (task1):} R2$=$\asm{0x8786\_8584}, R3$=$\asm{0x7B7A\_7978}.\\
\textbf{Ans (task2):} R0$=$\asm{0x2000\_0008}, R3$=$\asm{0x8786\_8584}.\\
\textbf{Mem @0x30} after task1: \asm{80 81 82 83 / 84 85 86 87}? -- \emph{No}: task1 stored its OWN r2,r3 at p2$=$0x20, not p3. The 0x30 region is written by task2: \asm{80 81 82 83 / 84 85 86 87}.\\
\textbf{Prototypes:} \cee{void task1(int *p1,int *p2);} and \cee{void task2(int *p1,int *p2,int n);}.\\
\textbf{C of task1:} \cee{*p2 = *++p1; *(p2+1) = *++p1;}\\
\emph{Trap:} scaled-reg with negative R2 walks \emph{backward}: \asm{[R0,R2,LSL\#2]} with R2$=-1$ subtracts 4 from R0. Two-arg \asm{ldr [Rn],\#8} post-increments Rn.
\end{qp}

% =============================================================================
\subsection{Practice qz-25b qz6a}

\begin{qp}\textbf{P1 (45pt) -- LDRD/STRD + sign-extend bytes.}
\cee{uint8\_t ipt[20]} with \cee{ipt[i]=0x70+(i<<2)} $\Rightarrow$ \asm{70 74 78 7C 80 84 88 8C 90 94 98 9C A0 A4 A8 AC B0 B4 B8 BC}.
\begin{lstlisting}
task_a:                       @ R0=ipt, R1=opt
    ldrd r2,r3,[r0,#8]!       @ R0+=8 -> ipt+8
                              @ R2 = word@(ipt+8) = 0x9C988_4?
                              @ Actually: 0x9C98_9490
                              @ R3 = word@(ipt+12)= 0xACA8_A4A0
    strd r2,r3,[r1,#0]        @ opt[0]=R2, opt[1]=R3
    ldrsb r2,[r0,#4]!         @ R0 -> ipt+12
                              @ byte=0xA0 -> R2=0xFFFF_FFA0
    ldrsb r3,[r0,#4]!         @ R0 -> ipt+16
                              @ byte=0xB0 -> R3=0xFFFF_FFB0
    strd r2,r3,[r1,#8]        @ opt[2]=R2, opt[3]=R3 (no wb)
    bx   lr
\end{lstlisting}
\textbf{Printout:}
\begin{lstlisting}
opt[0]=0x9C989490, opt[1]=0xACA8A4A0,
opt[2]=0xFFFFFFA0, opt[3]=0xFFFFFFB0
\end{lstlisting}
\textbf{Final regs:} R0$=$\cee{ipt}$+16$, R1$=$\cee{opt} (no \asm{!} on last \asm{strd}).\\
\textbf{C translation:}
\begin{lstlisting}
void task_a(uint8_t *p, int32_t *q){
  p += 8;
  q[0] = *(int32_t*)p;
  q[1] = *(int32_t*)(p+4);
  p += 4;
  q[2] = (int32_t)(int8_t)*p;
  p += 4;
  q[3] = (int32_t)(int8_t)*p;
}
\end{lstlisting}
\emph{Trap:} \asm{LDRSB} on byte $\geq$\asm{0x80} gives \asm{0xFFFF\_FFxx}; on byte $<$\asm{0x80} gives \asm{0x0000\_00xx}. Cast chain \cee{(int32\_t)(int8\_t)} forces sign-extend in C.
\end{qp}

\begin{qp}\textbf{P2 (20pt) -- saturating clamp $[-5,+5]$.}
\begin{lstlisting}
int math_fun_c(int x){
  if (x < -5) x = -5;
  else if (x > 5) x = 5;
  return x;
}
@ x -> r0 ; return r0
math_fun_s:
    cmp  r0, #-5
    bge  math_fun_chk_hi    @ x>=-5: maybe still > 5
    mov  r0, #-5            @ x<-5
    bx   lr
math_fun_chk_hi:
    cmp  r0, #5
    ble  math_fun_end       @ -5<=x<=5: pass through
    mov  r0, #5             @ x>5
math_fun_end:
    bx   lr
\end{lstlisting}
\emph{Trap:} the C \cee{else if} means the second test runs ONLY when the first fell through. Don't put both clamps under unconditional branches -- otherwise you'd clamp $-10$ first to $-5$ then re-test and accidentally pass through. The \asm{bx lr} inside the low-clamp avoids re-running the high-clamp test.
\end{qp}

\begin{qp}\textbf{P3 (35pt) -- sum of negative elems of array.}
\begin{lstlisting}
int mp_neg_sum_array_c(int *arr, int N){
    int s = 0;
    for (int i = 0; i < N; i++)
        if (arr[i] < 0) s += arr[i];
    return s;
}

@ arr->r0, N->r1 ; return r0
@ tmp:r2=elem, r3=sum, r12=i (or reuse r1)
mp_neg_sum_array_s:
    mov   r3, #0             @ sum = 0
    mov   r12, #0            @ i   = 0
.Lloop:
    cmp   r12, r1
    bge   .Ldone             @ i>=N done (signed)
    ldr   r2, [r0, r12, lsl #2]   @ arr[i]
    cmp   r2, #0
    bge   .Lskip             @ arr[i]>=0 skip
    add   r3, r3, r2         @ s += arr[i]
.Lskip:
    add   r12, r12, #1
    b     .Lloop
.Ldone:
    mov   r0, r3
    bx    lr
\end{lstlisting}
\emph{Trap:} use \asm{BGE}/\asm{BLT} (signed) since \cee{int}; arr indexed via scaled reg \asm{LSL \#2} for \cee{int*}. R12 (IP) is caller-saved so no push needed; if you used R4--R11 you'd need \asm{push \{r4,lr\}}/\asm{pop \{r4,pc\}} for AAPCS compliance.
\end{qp}

% =============================================================================
\subsection{Carry-over patterns to memorize}
\begin{enumerate}
\item \textbf{Sign-ext loads}: \asm{LDRSB}/\asm{LDRSH} fill upper bits with the loaded MSB. Positive byte/HW $\to$ upper zero; negative $\to$ upper \asm{0xFF...}.
\item \textbf{Pre vs post writeback}: \asm{[Rn,\#i]!} updates Rn first, then loads from new Rn; \asm{[Rn],\#i} loads from old Rn, then advances. Both modify Rn.
\item \textbf{Scaled-register addressing}: \asm{LDR Rt,[Rn,Rm,LSL \#n]} where n$=\log_2$sizeof. Index in elements; result in bytes.
\item \textbf{LDRD/STRD pairing}: Rt$=$lower addr, Rt2$=$upper. Pair must be consecutive (Rt2$=$Rt$+$1), Rt typically even. \asm{[Rn,\#i]!} writes back to Rn; trailing \asm{,\#i} (no \asm{!}) does not.
\item \textbf{C ptr idioms}: \cee{*p++}=post-idx; \cee{*++p}=pre-idx; \cee{++*p}=load+ADD\#1+\textbf{store-back}; \cee{(*p)++}=load+copy+ADD on copy+store original.
\item \textbf{LE bytes in store}: R[7:0]$\to$lowest byte at addr; R[31:24]$\to$highest. STRH only writes low 16; STRB only low 8.
\item \textbf{Cond-branch \cee{||}}: 1st test true $\to$ jump to then; 1st false $\to$ test 2nd. \cee{\&\&}: 1st false $\to$ jump to else; 1st true $\to$ test 2nd.
\item \textbf{Pointer offset units}: \cee{*(p+k)} compiles to \asm{[Rn,\#k*sizeof(*p)]}: \cee{int*}$\to$\asm{\#4k}; \cee{int16\_t*}$\to$\asm{\#2k}; \cee{char*}$\to$\asm{\#k}.
\item \textbf{Signed vs unsigned div by $2^n$}: \asm{ASR \#n} (signed); \asm{LSR \#n} (unsigned). Match shift family to operand domain (and CCS family: G/L signed, H/L unsigned).
\item \textbf{Post-idx advance $\neq$ load size}: \asm{ldrsh r3,[r0],\#4} reads 2B but advances R0 by 4. The post-idx imm is independent of the load width.
\end{enumerate}

% ==============================================================================
% Common footer for both parts. Closes multicols + document.
% ==============================================================================
\end{multicols*}
\end{document}
