% Source: ME19_v1.html

\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{siunitx}
\usepackage{enumitem}

\title{ME19: Heat Transfer}
\author{}
\date{}

\begin{document}
\maketitle

\section*{ME19: Heat Transfer}

\begin{enumerate}[leftmargin=*,label=\textbf{Question \arabic*.}]

\item \hfill \textit{10 pts}

A system of gas decreases in volume from 3.2 m$^{3}$ to 1.1 m$^{3}$ at a constant pressure of 197 kPa. Calculate the work in kJ done \textbf{on}the gas during this process.



\emph{W}$_{on}$ = \rule{2.5cm}{0.4pt} kJ

\textit{Correct answer:} 413.7

\item \hfill \textit{10 pts}

In the given \emph{pV}-diagram (not to scale), \emph{p}$_{1}$ = 91 kPa, \emph{p}$_{2}$ = 212 kPa; \emph{V}$_{1}$ = 0.9 m$^{3}$, \emph{V}$_{2}$ = 2 m$^{3}$, and \emph{V}$_{3}$ = 3 m$^{3}$. Calculate the work in kJ done \textbf{by} the system during these processes.

 [Image: ME19\_PV\_step\_down.png]



\emph{W}= \rule{2.5cm}{0.4pt} kJ

\textit{Correct answer:} 324.2

\item \hfill \textit{10 pts}

In the given \emph{pV}-diagram (not to scale), \emph{p}$_{1}$ = 172 kPa, \emph{p}$_{2}$ = 415 kPa; \emph{V}$_{1}$ = 1.3 m$^{3}$, \emph{V}$_{2}$ = 4.5 m$^{3}$, and \emph{V}$_{3}$ = 5.1 m$^{3}$. Calculate the work in kJ done \textbf{on} the system during these processes.

    [Image: ME19\_PV\_5.png]



\emph{W}$_{on}$\emph{}= \rule{2.5cm}{0.4pt} kJ

\textit{Correct answer:} -1,115.3

\item \hfill \textit{10 pts}

Suppose a system of gas absorbs 644 J of energy while doing 0.62 times that amount of work on the environment. Calculate the change in the internal energy of the gas.

\textit{Correct answer:} 244.72

\item \hfill \textit{10 pts}

An ideal monatomic gas increases in pressure from 116 kPa to 208 kPa at a constant volume of 1.1 m$^{3}$. Calculate the energy in kJ transferred by heat during the process.

\textit{Correct answer:} 151.8

\item \hfill \textit{10 pts}

A monatomic ideal gas at constant volume of 1.4 m$^{3}$ increases pressure from 61 kPa to 110 kPa. Calculate the work done by the gas in kJ during this process.

\textit{Correct answer:} 0

\item \hfill \textit{10 pts}

A diatomic ideal gas contracts at constant pressure of 225 kPa from 3.9 m$^{3}$ to 1.7 m$^{3}$. Calculate the change in the internal energy in kJ during the process.

\textit{Correct answer:} -1,237.5

\item \hfill \textit{10 pts}

A monatomic ideal gas expands at constant pressure of 130 kPa from 1.2 m$^{3}$ to 4 m$^{3}$. Calculate the work done \textbf{by} the gas in kJ.

\textit{Correct answer:} 364

\item \hfill \textit{10 pts}

A monatomic ideal gas expands at constant temperature from 1.3 m$^{3}$ to 2 m$^{3}$. If the initial pressure is 159 kPa, calculate the work in kJ done \textbf{on} the gas during this process.



\emph{W}$_{on}$ = \rule{2.5cm}{0.4pt} kJ

\textit{Correct answer:} -89.0428

\item \hfill \textit{10 pts}

A monatomic ideal gas undergoes an isothermal compression from a volume of 3 m$^{3}$ and pressure 120 kPa to a final volume of 1.2 m$^{3}$. Calculate the work in kJ done \textbf{by} the gas on its environment.

\textit{Correct answer:} -329.8647

\item \hfill \textit{10 pts}

A diatomic ideal gas is compressed adiabatically from 2.8 m$^{3}$ to 1.4 m$^{3}$. If the initial pressure is 81 kPa, calculate the final pressure in kPa.

\textit{Correct answer:} 213.7603

\item \hfill \textit{10 pts}

A monatomic ideal gas increases in pressure adiabatically from an initial pressure of 1.00 × 10$^{5}$ Pa and volume of 3.00 m$^{3}$ to a final volume of 1.00 m$^{3}$.  Two answers are required.

(a) Calculate the final pressure. Answer:

    [ Select ]

  ["538 kPa", "567 kPa", "624 kPa", "599kPa", "4.09 kPa"]

(b) Calculate the change in the internal energy. Answer:

    [ Select ]

  ["486 kJ", "935 kJ", "589 kJ", "284 kJ", "777 kJ"]

\begin{itemize}
  \item 538 kPa
  \item 567 kPa
  \item 624 kPa
  \item 599kPa
  \item 4.09 kPa
  \item 486 kJ
  \item 935 kJ
  \item 589 kJ
  \item 284 kJ
  \item 777 kJ
\end{itemize}

\item \hfill \textit{10 pts}

A monatomic ideal gas expands adiabatically from 1.4 m$^{3}$ to 2.6 m$^{3}$. If the initial pressure is 304 kPa, calculate the change in the internal energy in kJ.

\textit{Correct answer:} -215.865

\item \hfill \textit{10 pts}

On a \emph{pV}-diagram, a diatomic gas process proceeds along the straight line from (3 m$^{3}$, 146 kPa) to ( 5 m$^{3}$, 541 kPa). Calculate the work done by the gas in kJ.

\textit{Correct answer:} 687

\item \hfill \textit{10 pts}

Choose the best answer at the end of each statement about ideal gas processes.

\textbf{(a)}In an isochoric process, the change in the internal energy equals the energy transfer.

    [ Select ]

  ["true", "false", "more information required"]

\textbf{(b)} In an isothermal process, the work done \textbf{by} the system equals the energy transfer by heat.

    [ Select ]

  ["False", "More information required", "True"]

\textbf{(c)} In an isobaric process, the change in the internal energy equals the energy transfer by heat minus the work done \textbf{on} the system.

    [ Select ]

  ["More information required", "true", "False"]

\textbf{(d)} In an adiabatic process, the work done \textbf{on} the system always equals the change in the internal energy.

    [ Select ]

  ["More information required", "False", "True"]

\begin{itemize}
  \item true
  \item false
  \item more information required
  \item False
  \item More information required
  \item True
  \item More information required
  \item true
  \item False
  \item More information required
  \item False
  \item True
\end{itemize}

\item \hfill \textit{10 pts}

Name the following thermodynamic processes by using one of the words below in each blank:

\textbf{(a)} The change in internal energy is zero:
.

\textbf{(b)} The work done by the system is zero:
.

\textbf{(c)}The energy transfer by heat is zero:
.

\textbf{Word list: adiabatic; isobaric; isothermal; isochoric; generic}

\begin{itemize}
  \item isothermal
  \item isochoric
  \item adiabatic
\end{itemize}

\end{enumerate}

\end{document}
