% Source: MQ16b_v1.html
\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{siunitx}
\usepackage{enumitem}

\title{MQ16b: Sound}
\author{}
\date{}

\begin{document}
\maketitle

\section*{MQ16b: Sound}

\begin{enumerate}[leftmargin=*,label=\textbf{Question \arabic*.}]
\item \hfill \textit{1 pts}

A pipe open at both ends has an \emph{n} = 2 harmonic of 506 Hz. Find the pipe's length in meters if the speed of sound is 344 m/s.

\item \hfill \textit{1 pts}

A pipe closed at one end has a length of 1 m. If the temperature of the room is 8°C, calculate the fundamental frequency in Hz.

Take the speed of sound to be $v=(331\;\mathrm{m/s})\sqrt{\frac{T}{273\;\mathrm{K}}}$ .

\item \hfill \textit{1 pts}

A motorcyclist is traveling at 31 m/s away from a police car at rest. If the policeman blows his horn at a frequency of 532 Hz, what frequency in Hz does the motorcyclist hear? The speed of sound is 347 m/s.

\item \hfill \textit{1 pts}

A bicyclist is traveling at 8 m/s towards a car at rest. If the driver blows her horn at a frequency of 498 Hz, what frequency in Hz does the bicyclist hear? The speed of sound is 340 m/s.

\item \hfill \textit{1 pts}

A car is traveling 15 m/s due east relative the ground. An ambulance is traveling on the same road, also going east at 34 m/s and away from the car, having already passed it. If the ambulance siren sounds at 1,483 Hz, what frequency does the driver of the car hear? The speed of sound in air: 340 m/s

\item \hfill \textit{1 pts}

A motorcyclist is traveling 30 m/s due west relative the ground. A fire truck is traveling on the same road, going east at 42 m/s and away from the motorcyclist, having already passed it going the opposite direction. If the fire truck siren sounds at 1,211 Hz, what frequency does the motorcyclist hear? Sound speed in air: 343 m/s.

\item \hfill \textit{1 pts}

Suppose fluid B has 4 times the bulk modulus and 2.2 times the density of fluid A. Find the ratio of the speed of sound in fluid B over the speed of sound in fluid A.

\item \hfill \textit{1 pts}

An organ pipe open at both ends creates a sound with fundamental frequency of 596 Hz. What is the fundamental frequency when a stopper comes down and closes the pipe at one end?

\end{enumerate}

\end{document}
