% Source: MQ17b_v1.html

\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{siunitx}
\usepackage{enumitem}

\title{MQ17b: Temperature and Thermal Expansion}
\author{}
\date{}

\begin{document}
\maketitle

\section*{MQ17b: Temperature and Thermal Expansion}

\begin{enumerate}[leftmargin=*,label=\textbf{Question \arabic*.}]

\item \hfill \textit{1 pts}

An insulated container holds 0.72 kg of water at 10°C. Calculate the mass of 90°C water that must be added so that the final equilibrium temperature is 35°C.

[The specific heat of water is 4190 J/(kg⋅K).]

\item \hfill \textit{1 pts}

Suppose a block of a substance has mass 8 kg and is in solid form at its melting point of -6°C. How much energy in kJ must be transferred to the block in order to put it in its liquid form at the same temperature? The latent heat of fusion of the substance is \emph{L}$_{f}$ = 160,829 J/kg.



\emph{Q} = \rule{2.5cm}{0.4pt} kJ

\item \hfill \textit{1 pts}

Suppose 4.5 kg of ice at 0°C is placed in 11.5 kg of water at 27°C. Calculate the final equilibrium temperature in Celsius. \emph{L}$_{f}$ = 3.35 x 10$^{5}$ J/kg; \emph{c} = 4190 J/(kg⋅K).

[Hint: There are two possible outcomes - the ice will be either fully or partially melted. Start by figuring out which occurs in this case.]

\item \hfill \textit{1 pts}

A window has dimensions 1.5 m × 2 m × 13.4 mm where the latter quantity is the thickness in millimeters. The temperature outside is -10°C, and inside it is 14°C. If the thermal conductivity of the glass is 0.8  W/(m⋅K), calculate the magnitude of the energy transfer rate through the window in kW.



\emph{H}= \rule{2.5cm}{0.4pt} kW

\item \hfill \textit{1 pts}

An iron ball has an area of 0.68 m$^{2}$. Suppose the temperature in the room is 23°C and the emissivity of the ball's surface is 0.57. Calculate the temperature of the ball in Celsius, if it is radiating energy at a net rate of 2,685 W.\textbf{}



\emph{T} = \rule{2.5cm}{0.4pt} °C

\end{enumerate}

\end{document}
