% Source: MQ19a_v1.html

\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{siunitx}
\usepackage{enumitem}

\title{MQ19a: Heat Transfer}
\author{}
\date{}

\begin{document}
\maketitle

\section*{MQ19a: Heat Transfer}

\begin{enumerate}[leftmargin=*,label=\textbf{Question \arabic*.}]

\item \hfill \textit{1 pts}

An ideal gas increases in pressure by 59 kPa at a constant volume of 7.7 m$^{3}$.  Calculate the work in kJ done \textbf{by} the gas.



\emph{W} = \rule{2.5cm}{0.4pt} kJ

\item \hfill \textit{1 pts}

In the given \emph{pV}-diagram (not to scale), \emph{p}$_{1}$ = 110 kPa, \emph{p}$_{2}$ = 399 kPa; \emph{V}$_{1}$ = 1.9 m$^{3}$, \emph{V}$_{2}$ = 4.3 m$^{3}$, and \emph{V}$_{3}$ = 5.1 m$^{3}$. Calculate the work in kJ done \textbf{on} the system during these processes.

    [Image: ME19\_PV\_5.png]



\emph{W}$_{on}$\emph{}= \rule{2.5cm}{0.4pt} kJ

\item \hfill \textit{1 pts}

In the given \emph{pV}-diagram (not to scale), \emph{p}$_{1}$ = 92.5 kPa, \emph{p}$_{2}$ = 263.9 kPa, \emph{p}$_{3}$ = 328.5 kPa; \emph{V}$_{1}$ = 1.1 m$^{3}$, \emph{V}$_{2}$ = 2.2 m$^{3}$, and \emph{V}$_{3}$ = 3.5 m$^{3}$. Calculate the work in kJ done \textbf{on} the system during these processes.

     [Image: ME19\_PV\_4.png]



\emph{W}$_{on}$\emph{}= \rule{2.5cm}{0.4pt} kJ

\item \hfill \textit{1 pts}

Suppose a system of gas absorbs 602 J of energy while doing 0.35 times that amount of work on the environment. Calculate the change in the internal energy of the gas.

\item \hfill \textit{1 pts}

A monatomic ideal gas at constant volume of 0.5 m$^{3}$ increases pressure from 98 kPa to 166 kPa. Calculate the change in the internal energy in kJ during this process.

\item \hfill \textit{1 pts}

A monatomic ideal gas at constant volume of 0.7 m$^{3}$ increases pressure from 74 kPa to 114 kPa. Calculate the work done by the gas in kJ during this process.

\item \hfill \textit{1 pts}

A monatomic ideal gas expands at constant pressure of 92 kPa from 1.6 m$^{3}$ to 3.8 m$^{3}$. Calculate the energy in kJ transferred by heat during the process.

\item \hfill \textit{1 pts}

A system of diatomic ideal gas contracts from a volume of 3.5 m$^{3}$ to a final volume of 1.4 m$^{3}$. If the process proceeds at constant pressure of 182 kPa, calculate the energy in kJ transferred by heat, where positive is energy transferred into the system, and negative is energy transferred out of the system.

\end{enumerate}

\end{document}
